### 2015-12-08 paper/paper.tex
Add paper stuff

\documentclass[10pt,twocolumn]{article} 
\usepackage{simpleConference}
\usepackage{amsmath}
\usepackage{times}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url,hyperref}

\begin{document}

\title{Extending Morphological Chains with Supervised Learning and Cross-Language Features}

\author{Calvin Huang (calvinh@mit.edu)\\
    Brian Shimanuki (bshimanuki@mit.edu)\\
    Charlotte Chen (czchen@mit.edu)
\\
}

\maketitle
\thispagestyle{empty}

\begin{abstract}
    Building on an existing algorithm to perform morphological analysis,
    based on using semantic features and contrastive estimation to detect morphological chains,
    we investigate the effects of various types of extensions to this algorithm.
    We began by adding certain features to the model and heuristics for prefix/suffix selection.
    In addition, we extend the unsupervised model with the log-likelihood of known word segmentations,
    producing a semi-supervised model.
    Finally, we also attempt to use Turkish morphological parents and an English-Turkish translation
    dictionary to help detect irregular English word segmentations and transformations.
\end{abstract}

% TODO explain existing code

\section{Semi-supervised Learning with Known Word Segmentations}
From known word segmentations, we can use these segmentations to reconstruct
the morphological chains that would lead to the correct segmentation of the word.

In addition to the neighborhood log-likelihood that we maximize,
we also want to maximize the log-likelihood of the correct segmentation of the training data.
If a given word can be expressed as a morphological chain, the likelihood of its correct segmentation
is equal to the product of the likelihoods of each word-to-parent transition within the morphological chain.

Therefore, the likelihood of the segmentation of the training data is the product of the likelihoods
of each correct word-to-parent transition for every morphological chain in the training data.

Therefore, the likelihood of the training data is given as such:
\begin{equation}
    \prod_{w, c \in D} P(c | w) = \prod_{w, c \in D}\left(\prod_{w', z' \in MC(w)} P(z | w)\right)
\end{equation}

explanation blah
%TODO better explanation

Therefore, the log-likelihood of our training data can be expressed as
\begin{equation}
    \begin{split}
        L_S(\theta; D) &= \log{\prod_{w, z \in D} P(z|w)} \\
                     &= \sum_{w, z \in D} \left(\theta \cdot \phi(w, z) - \log\left(\sum_{z' \in C(w)} e^{\theta \cdot \phi(w, z')} \right)\right)
\end{split}
\end{equation}
where $w, z$ represent every morphological chain transition implied in the training data.

We can also express its gradient as such:
\begin{equation}
    \begin{split}
        \nabla_\theta L_S &= \sum_{w, z \in D} \left(\phi(w, z) - \frac{\sum_{z' \in C(w)} \phi(w, z') e^{\theta \cdot \phi(w, z')} }{\sum_{z' \in C(w)} e^{\theta \cdot \phi(w, z')}}\right)
    \end{split}
\end{equation}

Adding this to our original log-likelihood from contrastive estimation
along with an L2 regularization term, we get our objective:
\begin{equation}
    \begin{split}
        L(\theta; D) &= L_{CE}(\theta; D) + \alpha L_S(\theta; D) - \lambda \| \theta \|^2 \\
        \nabla_\theta L &= \nabla_\theta L_{CE} + \alpha \nabla_\theta L_S - 2\lambda \theta
    \end{split}
\end{equation}
where $\alpha$ is a free parameter that specifies by how much to weight the labeled training data likelihood
against the contrastive estimation likelihood.
As before, with its gradient given above, we minimize $L(\theta, D)$ with LBFGS-B.

\section{Using \LaTeX with PDF Figures}

This is a sample document for use with pdflatex, which is
a program that is included with the Miktex distribution
that directly produces PDF files from \LaTeX sources.
To run \LaTeX on this file, you need the following files:
\begin{enumerate}
\item templatePDF.tex (this file)
\item figure.pdf (the figure file)
\item simpleConference.sty (style file)
\item refs.bib (bibiliography file)
\end{enumerate}
\noindent
To create a PDF file, execute the following commands:
\begin{enumerate}
\item pdflatex templatePDF
\item bibtex templatePDF
\item pdflatex templatePDF
\item pdflatex templatePDF
\end{enumerate}
\noindent
Yes (strangely) it is necessary to run pdflatex three times.
The result will be a PDF file (plus several other files that \LaTeX
produces).  You will need a mechanism, of course, for executing
commands on the command line. If you are using Windows, I recommend
installing Cygwin and using its bash shell.

\section{How to Include Vergil Diagrams as Figures}

Suppose you wish to include a figure, like that in figure \ref{fig-label}.
The simplest mechanism is to install Adobe Acrobat, which includes
a ``printer'' called ``Acrobat Distiller.'' Printing to this printer
creates a PDF file, which can be included in a document as shown
here.  To include Ptolemy II models \cite{PtolemyVol1:04},
just print to the distiller from within Vergil and reference
the PDF file in your \LaTeX document.

There is a bit more work to do, however.
The file that is produced by the distiller represents
a complete page, not the individual figure.
You can open it in using Acrobat (version 5.0 or later),
and select Document $\rightarrow$ Crop Pages from the menu.
In the resulting dialog, check ``Remove White Margins.''
Save the modified PDF file in a file and then reference
it in the \LaTeX file as shown in this example.

An alternative is to generate EPS (encapsulated postscript),
but the process is much more complex and fragile.
I recommend using pdflatex and Adobe Acrobat.

\bibliographystyle{abbrv}
\bibliography{refs}
\end{document}

### 2015-12-08 pysrc/morphochain.py
gold

@@ -351,24 +351,38 @@ class MorphoChain(object):
         def chain(segs, tags):
             def getAffix(s, t):
                 return s if t[0] == '+' else t[:t.find('_')]
+            def suffix_type(s, t):
+                if t[0] == '+':
+                    return ParentType.SUFFIX
+                t = t[:t.find('_')]
+                if s == t:
+                    return ParentType.SUFFIX
+                elif len(s) == len(t):
+                    return ParentType.MODIFY
+                elif s[:-1] == t:
+                    return ParentType.REPEAT
+                else:
+                    return ParentType.DELETE
             if len(segs) <= 1:
-                return getAffix(segs[0], tags[0])
+                root = getAffix(segs[0], tags[0])
+                return [(root, ParentTransformation(root, ParentType.STOP))]
             word = ''.join(segs[:-1]) + getAffix(segs[-1], tags[-1])
             word = word.replace('~', '') # null segments
             parent_suf = ''.join(segs[:-2]) + getAffix(segs[-2], tags[-2])
             parent_suf = parent_suf.replace('~', '')
-            pre_parent = ''.join(segs[1:])
+            pre_parent = ''.join(segs[1:-1]) + getAffix(segs[-1], tags[-1])
             pre_parent = pre_parent.replace('~', '')
-            if segs[-1] == '~':
-                return [word] + chain(segs[:-1], tags[:-1])
-            elif tags[-1][0] == '+': # inflectional
-                return [word] + chain(segs[:-1], tags[:-1])
+            type_suf = suffix_type(segs[-2], tags[-2])
+            pre_pair = [(word, ParentTransformation(pre_parent, ParentType.PREFIX))]
+            pair_suf = [(word, ParentTransformation(parent_suf, type_suf))]
+            if tags[-1][0] == '+': # inflectional
+                return pair_suf + chain(segs[:-1], tags[:-1])
             score_suf = parentScore(parent_suf, word)
             pre_score = parentScore(pre_parent, word)
-            if pre_score < score_suf:
-                return [word] + chain(segs[1:], tags[1:])
+            if pre_score > score_suf:
+                return pre_pair + chain(segs[1:], tags[1:])
             else:
-                return [word] + chain(segs[:-1], tags[:-1])
+                return pair_suf + chain(segs[:-1], tags[:-1])
         d = {}
         for word, (g_segs, g_tags) in segmentations.iteritems():
             d[word] = []

### 2015-12-08 paper/paper.tex

@@
 \section{Semi-supervised Learning with Known Word Segmentations}
-From known word segmentations, we can use these segmentations to reconstruct
+In addition to maximizing the log-likelihood of the known wordlist,
+we can also use explicit feedback regarding the correct segmentation of known words.
+From these known segmentations, we can use these segmentations to reconstruct
 the morphological chains that would lead to the correct segmentation of the word.
@@
 As before, with its gradient given above, we minimize $L(\theta, D)$ with LBFGS-B.
+
+\section{Using Cross-Language Features and Candidates}
+Due to the simple nature of our parent candidate generation routine,
+there are many cases in which the potential parent is not actually generated;
+for example, the morphological parent of "feet" is the singular "foot";
+however, there is no prefix or suffix that turns one into the other, and
+therefore the parent for "feet" is not generated by our system.
+
+However, other languages may not have such an irregular transformation
+for the same words, for example in Turkish the corresponding translations
+for "feet" and "foot" are "ayaklar" and "ayak". A properly trained Turkish
+morphological parent model will correctly detect "ayak" as a parent of "ayaklar",
+and can be used to suggest morphological parents in English after translation.
+
+Since there are multiple possible morphological parents for a given word, and multiple
+possible translations in either direction, this leads to many possible candidates for a given word,
+we generate the set of possible parent candidates (from translation)
+\begin{equation}
+    C_2(w) = \bigcup_{w_t \in T_{ET}(w)}\left(\bigcup_{w_{tp} \in MP_T (w_t)} T_{TE}(w_{tp})\right)
+\end{equation}
+where $T_{ET}$ returns the set of possible Turkish translations
+for an english word; $MP_T$ returns the set of possible parents for some Turkish word,
+$T_{TE}$ returns the set of possible english translations for an english word,
+and this potentially large set of candidates is pruned with heuristics judging
+how similar the candidate is to the original word.
 

### 2015-12-08 pysrc/morphochain.py
gold segs score

@@ -283,7 +283,6 @@ class MorphoChain(object):
 
     def similarity(self, w1, w2):
         if w1 not in self.wordvectors or w2 not in self.wordvectors:
-            # is this what we want to return? or just 0?
             return 0.0
         return self.wordvectors.similarity(w1, w2)
 
@@ -343,10 +342,19 @@ class MorphoChain(object):
         if segmentations is None:
             segmentations = self.segmentations
         def parentScore(parent, word): # can we train on these weights :)
+            FREQ_WEIGHT = 1
+            SIM_WEIGHT = 1
+            SIM_BASE = 12
+            LEN_WEIGHT = 1
+            LEN_BASE = 5
+            LENGTH_WEIGHT = 30
+            LENGTH_POWER = 2
             score = 0
             if parent in self.vocab:
-                score += 1
-            score += max(0, self.similarity(parent, word))
+                score += FREQ_WEIGHT * math.log(self.vocab[parent])
+            score += SIM_WEIGHT * SIM_BASE ** max(0, self.similarity(parent, word))
+            score += LEN_WEIGHT * LEN_BASE ** (float(len(parent)) / len(word))
+            score -= LENGTH_WEIGHT / max(1, len(parent)) ** LENGTH_POWER
             return score
         def chain(segs, tags):
             def getAffix(s, t):

### 2015-12-08 paper/paper.tex
add translation dicts

@@
 possible translations in either direction, this leads to many possible candidates for a given word,
-we generate the set of possible parent candidates (from translation)
+we generate a set of possible parent candidates (from translation):
 \begin{equation}
@@
 and this potentially large set of candidates is pruned with heuristics judging
-how similar the candidate is to the original word.
+how similar the candidate is to the original word. We then extend our original parent candidate
+with this parent candidate set.
+
+Additionally, we extend our feature set to include whether or not each candidate was
+generated by a turkish translation (a strong indicator for parent-ness) and how confident
+the Turkish model was in selecting a parent.
 

### 2015-12-08 pysrc/fileio.py

@@ -1,5 +1,7 @@
 from gensim.models.word2vec import Word2Vec
-from collections import Counter
+from collections import Counter, defaultdict
+import re
+import codecs
 try:
     import cPickle as pickle
 except:
@@ -52,6 +54,29 @@ def read_words(filename):
         for line in f.readlines():
             s.add(line.strip())
     return s
+
+
+ROMAN_TT = str.maketrans('öüçýiþðâ', 'OUCiISGA')
+def romanize_turkish(s):
+    return s.lower().translate(ROMAN_TT)
+
+
+def read_dictionary(filename):
+    eng_to_tur = defaultdict(lambda: [])
+    tur_to_eng = defaultdict(lambda: [])
+
+    WWDR = re.compile(r'^([^ ]+)\t([^ ]+)\t')
+    with codecs.open(filename, 'r', 'latin-1') as f:
+        for line in f:
+            match = WWDR.match(line)
+            if match:
+                eng, tur = match.groups()
+                eng = eng.lower()
+                tur = romanize_turkish(tur)
+                eng_to_tur[eng].append(tur)
+                tur_to_eng[tur].append(eng)
+    return eng_to_tur, tur_to_eng
+
 
 # readCorpus('data/goldstd_trainset.segmentation.eng.txt', 'data/traincorpus.p')
 # readCorpus('data/goldstd_develset.segmentation.eng.txt', 'data/devcorpus.p')

### 2015-12-08 pysrc/__main__.py
Supervised up and running. Results below are from small-small. Also more stuff for turkish. Unicodefied and romanized the dictionary.
Training Set
Precision: 0.727701746107
Recall: 0.800207576544

Dev Set
Precision: 0.681415929204
Recall: 0.758908263836

@@ -12,11 +12,12 @@ except:
 readline.parse_and_bind('tab: complete')
 
 """
-args: [optimize | load | run] [now] [wordlist_size] [wordvector_size]
+args: [optimize | load | run] [supervised] [now] [wordlist_size] [wordvector_size]
 
 optimize: run the optimizer to find weights
 load: load weights to do stuff with
 run: like load, but has a simple text interface
+supervised: for the supervised model
 now: option for load and run to use dictvectorizer and weights currently being generated
 wordlist_size: default is smallw, can also set to medw, largew, or fullw for different sizes of training
 wordvector_size: default is smallv, can also set to medv or fullv for different sizes of word2vec
@@ -65,11 +66,19 @@ if __name__ == '__main__':
             pickle.dump(en_morpho.dictvectorizer, f)
         with open('out_py/dictvectorizer.%s-%s.p' % (w_size,v_size), 'wb') as f:
             pickle.dump(en_morpho.dictvectorizer, f)
-        print('training data saved, optimizing weights')
-        weights = optimize_weights(*train)
-        en_morpho.setWeightVector(weights)
-        with open('out_py/weights.%s-%s.p' % (w_size,v_size), 'wb') as f:
-            pickle.dump(weights, f)
+        if 'supervised' in sys.argv:
+            sups = en_morpho.genGoldsegTrainingData()
+            print('training data saved, optimizing weights')
+            weights = optimize_weights_supervised(*(train + sups))
+            en_morpho.setWeightVector(weights)
+            with open('out_py/weights.supervised.%s-%s.p' % (w_size,v_size), 'wb') as f:
+                pickle.dump(weights, f)
+        else:
+            print('training data saved, optimizing supervised weights')
+            weights = optimize_weights(*train)
+            en_morpho.setWeightVector(weights)
+            with open('out_py/weights.%s-%s.p' % (w_size,v_size), 'wb') as f:
+                pickle.dump(weights, f)
 
     elif 'load' in sys.argv or 'run' in sys.argv:
         def loadweights():
@@ -84,7 +93,7 @@ if __name__ == '__main__':
         else:
             with open('out_py/dictvectorizer.%s-%s.p' % (w_size,v_size), 'rb') as f:
                 en_kwargs['dictvectorizer'] = pickle.load(f)
-            with open('out_py/weights.%s-%s.p' % (w_size,v_size), 'rb') as f:
[…]

### 2015-12-08 pysrc/fileio.py

@@ -1,3 +1,5 @@
+# -*- coding: utf-8 -*-
+import sys
 from gensim.models.word2vec import Word2Vec
 from collections import Counter, defaultdict
 import re
@@ -56,7 +58,7 @@ def read_words(filename):
     return s
 
 
-ROMAN_TT = str.maketrans('öüçýiþðâ', 'OUCiISGA')
+ROMAN_TT = {ord(c): ord(t) for c, t in zip('öüçýiþðâ', 'OUCiISGA')}
 def romanize_turkish(s):
     return s.lower().translate(ROMAN_TT)
 

### 2015-12-08 pysrc/morphochain.py

@@ -1,6 +1,7 @@
 from collections import namedtuple, Counter
 import math
 import string
+import itertools
 from enum import Enum
 import numpy as np
 import scipy.spatial.distance
@@ -209,11 +210,13 @@ class MorphoChain(object):
         return X, nzs, widsneighbors
 
     def genGoldsegTrainingData(self):
-        # TODO get chains
         dicts = []
         nzs = []
         cxs = []
-        for word, parent in stuff:
+        chainsets = self.genGoldChains().itervalues()
+        chains = itertools.chain(*chainsets)
+        wordpairs = itertools.chain(*chains)
+        for word, parent in wordpairs:
             parentsFeatures = self.getParentsFeatures(word)
 
             parents = list(parentsFeatures.keys())
@@ -228,9 +231,6 @@ class MorphoChain(object):
 
         X = self.dictvectorizer.transform(dicts)
         return X, nzs, cxs
-
-
-
 
     def scoreFeatures(self, featureDict):
         fv = self.dictvectorizer.transform(featureDict)

### 2015-12-08 pysrc/objective.py

@@ -80,7 +80,8 @@ def get_logprob_fn(X, nzs, cxs):
         logprobs = Xp[cxs] - np.log(normcs)
         grad = Xdense[cxs].sum(0)
         for i, (a, b) in enumerate(idxs):
-            grad -= (Xdense[a:b] * Xpexp[a:b]) / normcs[i]
+            # check that this is what you want to do
+            grad -= np.dot(Xpexp[a:b].reshape(1,-1), Xdense[a:b]).flatten() / normcs[i]
         return -logprobs.sum(), -grad  # want to maximize
 
     return f

### 2015-12-08 tr2rom.sh

@@ -1,0 +1,12 @@
+tr2rom() {
+	sed -e 's/[âÂ]/A/g' \
+		-e 's/[çÇ]/C/g' \
+		-e 's/[ğĞÐð]/G/g' \
+		-e 's/[ıIİİîÎÝý]/I/g' \
+		-e 's/[öÖ]/O/g' \
+		-e 's/[şŞÞþ]/S/g' \
+		-e 's/[üÜûÛ]/U/g'
+}
+
+[ $# -ge 1 -a -f "$1" ] && input="$1" || input="-"
+cat $input | tr2rom

### 2015-12-08 pysrc/__main__.py
nicer arg parsing

@@ -3,6 +3,7 @@ import rlcompleter
 import readline
 import sys
 import os.path
+import argparse
 
 try:
     import cPickle as pickle
@@ -12,38 +13,39 @@ except:
 readline.parse_and_bind('tab: complete')
 
 """
-args: [optimize | load | run] [supervised] [now] [wordlist_size] [wordvector_size]
+args: [optimize | load | run] [--supervised=lamb] [--now]
+[--vectors=[small, med, large, full]] [--vocab=[small, med, large, full]]
 
 optimize: run the optimizer to find weights
 load: load weights to do stuff with
 run: like load, but has a simple text interface
-supervised: for the supervised model
+supervised: for the supervised model, if lamb is not 0, by how much to weight the likelihood
 now: option for load and run to use dictvectorizer and weights currently being generated
 wordlist_size: default is smallw, can also set to medw, largew, or fullw for different sizes of training
 wordvector_size: default is smallv, can also set to medv or fullv for different sizes of word2vec
 """
 
 if __name__ == '__main__':
-    w_size = 'small'
+    parser = argparse.ArgumentParser()
+    parser.add_argument('command', choices=['optimize', 'load', 'run'])
+    parser.add_argument('--now', action='store_true')
+    parser.add_argument('--vocab', choices=['small', 'med', 'large', 'full'], default='small')
+    parser.add_argument('--vectors', choices=['small', 'med', 'large', 'full'], default='small')
+    parser.add_argument('--supervised', type=float, default=0.0)
+    args = parser.parse_args()
     file_w = 'data/wordlist-2010.small.txt'
-    if 'medw' in sys.argv:
-        w_size = 'med'
+    if args.vocab == 'med':
         file_w = 'data/wordlist-2010.med.txt'
-    if 'largew' in sys.argv:
-        w_size = 'large'
+    if args.vocab == 'large':
         file_w = 'data/wordlist-2010.large.txt'
-    if 'fullw' in sys.argv:
-        w_size = 'full'
+    if args.vocab == 'full':
         file_w = 'data/wordlist-2010.eng.txt'
     en_wordcounts = read_wordcounts(file_w)
 
-    v_size = 'small'
     file_v = 'data/en-wordvectors200_small.txt'
-    if 'medv' in sys.argv:
-        v_size = 'med'
+    if args.vectors == 'med':
         file_v = 'data/en-wordvectors200_med.txt'
-    if 'fullv' in sys.argv:
-        v_size = 'full'
[…]

### 2015-12-08 pysrc/morphochain.py

@@ -213,7 +213,7 @@ class MorphoChain(object):
         dicts = []
         nzs = []
         cxs = []
-        chainsets = self.genGoldChains().itervalues()
+        chainsets = self.genGoldChains().values()
         chains = itertools.chain(*chainsets)
         wordpairs = itertools.chain(*chains)
         for word, parent in wordpairs:
@@ -392,7 +392,7 @@ class MorphoChain(object):
             else:
                 return pair_suf + chain(segs[:-1], tags[:-1])
         d = {}
-        for word, (g_segs, g_tags) in segmentations.iteritems():
+        for word, (g_segs, g_tags) in segmentations.items():
             d[word] = []
             for segs, tags in zip(g_segs, g_tags):
                 d[word].append(chain(segs, tags))

### 2015-12-08 pysrc/objective.py
decorator to properly show stats

@@ -11,6 +11,26 @@ except:
 BOUNDS = 50
 MAX_ITERS = 500
 
+def output_decorate(func):
+    iteration = [0]
+
+    def newf(weights):
+        stime = time.time()
+        fv, gv = func(weights)
+        iteration[0] += 1
+        with open('out_py/weights.p', 'wb') as f:
+            pickle.dump(weights, f)
+        print('Call %s' % iteration[0])
+        print('\tWeights range: %s %s' % (weights.min(), weights.max()))
+        print('\tWeights norm (d=%s): %s' % (weights.size, np.linalg.norm(weights)))
+        print('\tFunction: %s' % fv)
+        print('\tGradient range: %s %s' % (gv.min(), gv.max()))
+        print('\tTime: %s' % (time.time() - stime))
+
+        return fv, gv
+    return newf
+
+
 def optimize_weights(X, nzs, widsneighbors, lamb=0, output=True):
     '''
     X is a feature matrix, where each row corresponds
@@ -31,7 +51,9 @@ def optimize_weights(X, nzs, widsneighbors, lamb=0, output=True):
 
     widsneighbors contains tuples, each tuple contains (the ids of the words, ids of the word's neighbors).
     '''
-    f = get_optimizer_fn(X, nzs, widsneighbors, lamb, output)
+    f = get_optimizer_fn(X, nzs, widsneighbors, lamb)
+    if output:
+        f = output_decorate(f)
     return scipy.optimize.fmin_l_bfgs_b(
         f,
         np.zeros_like(X[0].toarray()).T,
@@ -44,12 +66,15 @@ def optimize_weights(X, nzs, widsneighbors, lamb=0, output=True):
 def optimize_weights_supervised(X, nzs, widsneighbors,
                                 Xsup, nzs_sup, cxs_sup,
                                 lamb=0, lamb2=1, output=True):
-    fus = get_optimizer_fn(X, nzs, widsneighbors, lamb, output)
+    fus = get_optimizer_fn(X, nzs, widsneighbors, lamb)
     fsup = get_logprob_fn(Xsup, nzs_sup, cxs_sup)
     def fcomb(weights):
         f1, g1 = fus(weights)
         f2, g2 = fsup(weights)
         return f1 + lamb2 * f2, g1 + lamb2 * g2
+
+    if output:
+        fcomb = output_decorate(fcomb)
 
     return scipy.optimize.fmin_l_bfgs_b(
         fcomb,
[…]

### 2015-12-08 pysrc/__main__.py
print out segmentations when optimizing too

@@ -100,7 +100,8 @@ if __name__ == '__main__':
                                                     if args.supervised else '', args.vocab, args.vectors), 'rb') as f:
                 en_kwargs['weightvector'] = pickle.load(f)
             en_morpho = MorphoChain(*en_args, **en_kwargs)
-        en_morpho.computeAccuracy(en_devsegmentations)
+
+    en_morpho.computeAccuracy(en_devsegmentations)
 
     if args.command == 'run':
         word = raw_input("Enter word: ")

### 2015-12-08 pysrc/__init__.py
make imports a little more sane (so can use ipython and stuff), use python -m to run now

@@ -1,6 +1,6 @@
-from fileio import *
-from morphochain import *
-from objective import *
+from pysrc.fileio import *
+from pysrc.morphochain import *
+from pysrc.objective import *
 import pprint
 
 with open('data/prefix_list.p', 'rb') as f:

### 2015-12-08 pysrc/__main__.py

@@ -1,4 +1,4 @@
-from __init__ import *
+from pysrc import *
 import rlcompleter
 import readline
 import sys

### 2015-12-08 pysrc/morphochain.py

@@ -8,7 +8,7 @@ import scipy.spatial.distance
 import sklearn.linear_model
 from sklearn.feature_extraction import DictVectorizer
 
-import accuracy
+import pysrc.accuracy
 
 ParentTransformation = namedtuple('ParentTransformation',
                                   ['parentword', 'transformtype'])

### 2015-12-08 pysrc/suffixes.py

@@ -1,4 +1,4 @@
-import fileio
+import pysrc.fileio
 from collections import Counter, defaultdict
 import math
 import numpy as np

### 2015-12-08 pysrc/__init__.py
oops, add language options

@@ -1,20 +1,56 @@
-from pysrc.fileio import *
-from pysrc.morphochain import *
-from pysrc.objective import *
+from pysrc import fileio, morphochain, objective
 import pprint
+import os
 
-with open('data/prefix_list.p', 'rb') as f:
-    en_prefixes = pickle.load(f)
-with open('data/suffix_list.p', 'rb') as f:
-    en_suffixes = pickle.load(f)
-with open('data/prefix_corr2.p', 'rb') as f:
-    en_prefixes_corr = pickle.load(f)
-with open('data/suffix_corr2.p', 'rb') as f:
-    en_suffixes_corr = pickle.load(f)
-en_affixes = (en_prefixes, en_suffixes)
-en_affix_corr = (en_prefixes_corr, en_suffixes_corr)
+try:
+    import cPickle as pickle
+except:
+    import pickle
 
-with open('data/traincorpus.p', 'rb') as f:
-    en_trainsegmentations = pickle.load(f)
-with open('data/devcorpus.p', 'rb') as f:
-    en_devsegmentations = pickle.load(f)
+
+def get_prefixes_affixes(lang='eng'):
+    assert lang == 'eng' or lang == 'tur'
+
+    with open('data/%s_prefix_list.p' % lang, 'rb') as f:
+        prefixes = pickle.load(f)
+    with open('data/%s_suffix_list.p' % lang, 'rb') as f:
+        suffixes = pickle.load(f)
+    with open('data/%s_prefix_corr2.p' % lang, 'rb') as f:
+        prefixes_corr = pickle.load(f)
+    with open('data/%s_suffix_corr2.p' % lang, 'rb') as f:
+        suffixes_corr = pickle.load(f)
+    affixes = (prefixes, suffixes)
+    affix_corr = (prefixes_corr, suffixes_corr)
+    return affixes, affix_corr
+
+
+def get_segmentations(lang='eng'):
+    assert lang == 'eng' or lang == 'tur'
+
+    with open('data/%s_traincorpus.p' % lang, 'rb') as f:
+        trainsegmentations = pickle.load(f)
+    with open('data/%s_devcorpus.p' % lang, 'rb') as f:
+        devsegmentations = pickle.load(f)
+    return trainsegmentations, devsegmentations
+
+
+def get_wordlist(size='small', lang='eng'):
+    if size == 'full':
+        fname = 'data/wordlist-2010.%s.txt' % lang
+    else:
[…]

### 2015-12-08 pysrc/__main__.py

@@ -1,4 +1,5 @@
 from pysrc import *
+from pysrc.morphochain import MorphoChain
 import rlcompleter
 import readline
 import sys
@@ -31,55 +32,38 @@ if __name__ == '__main__':
     parser.add_argument('--now', action='store_true')
     parser.add_argument('--vocab', choices=['small', 'med', 'large', 'full'], default='small')
     parser.add_argument('--vectors', choices=['small', 'med', 'large', 'full'], default='small')
+    parser.add_argument('--lang', choices=['eng', 'tur'], default='eng')
     parser.add_argument('--supervised', type=float, default=0.0)
     args = parser.parse_args()
-    file_w = 'data/wordlist-2010.small.txt'
-    if args.vocab == 'med':
-        file_w = 'data/wordlist-2010.med.txt'
-    if args.vocab == 'large':
-        file_w = 'data/wordlist-2010.large.txt'
-    if args.vocab == 'full':
-        file_w = 'data/wordlist-2010.eng.txt'
-    en_wordcounts = read_wordcounts(file_w)
-
-    file_v = 'data/en-wordvectors200_small.txt'
-    if args.vectors == 'med':
-        file_v = 'data/en-wordvectors200_med.txt'
-    if args.vectors == 'full':
-        file_v = 'data/en-wordvectors200_filtered.txt'
-
-    binfile_v = file_v[:-3] + 'bin'
-    if os.path.isfile(binfile_v):
-        en_wordvectors = load_wordvectors(binfile_v, binary=True)
-    else:
-        en_wordvectors = load_wordvectors(file_v)
 
     if sys.version_info >= (3, 0):
         raw_input = input  # ghettooooooooo
 
-    en_args = (en_wordvectors, en_wordcounts, en_affixes, en_affix_corr)
-    en_kwargs = {'segmentations': en_trainsegmentations}
+    trainsegmentations, devsegmentations = get_segmentations(args.lang)
+    mc_args = (get_wordvectors(args.vectors, args.lang), get_wordlist(args.vocab, args.lang),
+               *get_prefixes_affixes(args.lang))
+    mc_kwargs = {'segmentations': trainsegmentations}
 
     if args.command == 'optimize':
-        en_morpho = MorphoChain(*en_args, **en_kwargs)
+        morpho = MorphoChain(*mc_args, **mc_kwargs)
         print('generating training data')
-        train = en_morpho.genTrainingData()
+        train = morpho.genTrainingData()
[…]

### 2015-12-08 pysrc/morphochain.py

@@ -8,7 +8,7 @@ import scipy.spatial.distance
 import sklearn.linear_model
 from sklearn.feature_extraction import DictVectorizer
 
-import pysrc.accuracy
+from pysrc import accuracy
 
 ParentTransformation = namedtuple('ParentTransformation',
                                   ['parentword', 'transformtype'])

### 2015-12-09 pysrc/morphochain.py
kimchi

@@ -11,7 +11,7 @@ from sklearn.feature_extraction import DictVectorizer
 import accuracy
 
 ParentTransformation = namedtuple('ParentTransformation',
-                                  ['parentword', 'transformtype'])
+                                  ['parentword', 'transformtype', 'olangconfidence'])
 
 
 class ParentType():
@@ -21,13 +21,15 @@ class ParentType():
     MODIFY = 'MODIFY'
     DELETE = 'DELETE'
     REPEAT = 'REPEAT'
+    OLANG = 'OLANG'
     TOPNEIGHBOURS = 7
 
 
 class MorphoChain(object):
     def __init__(self, wordvectors, vocab, affixes, affixNeighbours, dictionary=None,
                  alphabet=string.ascii_lowercase, dictvectorizer=None,
-                 weightvector=None, segmentations=None):
+                 weightvector=None, segmentations=None, translations=None,
+                 secondLangChain=None):
         self.wordvectors = wordvectors
         self.vocab = vocab
         self.dictionary = dictionary
@@ -37,6 +39,8 @@ class MorphoChain(object):
         self.prefixNeighbours, self.suffixNeighbours = affixNeighbours
         self.weightvector = weightvector
         self.segmentations = segmentations
+        self.translations, self.invTranslations = translations
+        self.secondLangChain = secondLangChain
 
     def setWeightVector(self, weightvector):
         self.weightvector = weightvector
@@ -48,7 +52,7 @@ class MorphoChain(object):
         parentsAndFeatures = {}
         for z in self.genCandidates(w):
             parentsAndFeatures[z] = self.getFeatures(w, z)
-        z = ParentTransformation(w, ParentType.STOP)
+        z = ParentTransformation(w, ParentType.STOP, None)
         parentsAndFeatures[z] = self.getFeatures(
             w, z, max([d['cos'] for d in parentsAndFeatures.values()] or [0]))
         return parentsAndFeatures
@@ -135,10 +139,10 @@ class MorphoChain(object):
         for x in range(len(word) // 2, len(word)):
             parent = word[:x]
             if parent:
-                candidates.append(ParentTransformation(parent, ParentType.SUFFIX))
+                candidates.append(ParentTransformation(parent, ParentType.SUFFIX, None))
             #planning - plan - (n)ing
             if x > 0 and word[x] == word[x - 1]:
-                candidates.append(ParentTransformation(parent, ParentType.REPEAT))
+                candidates.append(ParentTransformation(parent, ParentType.REPEAT, None))
[…]

### 2015-12-09 pysrc/__main__.py
fix accuracy calculation

@@ -91,9 +91,9 @@ if __name__ == '__main__':
         word = raw_input("Enter word: ")
         while True:
             if word == 'ACCURACY_TRAIN':
-                print(morpho.computeAccuracy())
+                morpho.computeAccuracy()
             elif word =='ACCURACY_DEVEL':
-                print(morpho.computeAccuracy(devsegmentations))
+                morpho.computeAccuracy(devsegmentations)
             elif word == 'RELOAD' and args.now:
                 loadweights()
             elif word == 'EXIT':

### 2015-12-09 pysrc/accuracy.py

@@ -5,6 +5,10 @@ except:
     import pickle
 
 
+def seg_points(segs):
+    return set(np.cumsum([len(i) for i in segs])[:-1])
+
+
 def score(gold_segs_list, predictions_segs_list):
     correct = 0
     goldTotal = 0
@@ -16,12 +20,12 @@ def score(gold_segs_list, predictions_segs_list):
         if word not in gold_segs_list:
             print('Error: %s' % word)
             continue
-        predTotal += len(pred_segs)
+        predTotal += len(pred_segs) - 1
         bestc = 0
         bestg = 0
         for gold_segs in gold_segs_list[word][0]:
-            gset = set(i for i in gold_segs if i != '~')
-            pset = set(pred_segs)
+            gset = seg_points(gold_segs)
+            pset = seg_points(pred_segs)
             num_correct_segs = len(gset & pset)
             num_gold_segs = len(gold_segs)
             if (bestc < num_correct_segs) or (bestc == num_correct_segs
@@ -37,5 +41,6 @@ def score(gold_segs_list, predictions_segs_list):
         correct += bestc
     precision = float(correct) / predTotal
     recall = float(correct) / goldTotal
-    print('Precision: %s\nRecall: %s' % (precision, recall))
-    return precision, recall
+    f1 = 2 * precision * recall / (precision + recall)
+    print('Precision: %s\nRecall: %s\nF1: %s' % (precision, recall, f1))
+    return precision, recall, f1

### 2015-12-09 pysrc/__main__.py
syntax fixes (at least for py2), tur generation For the turkish prefixes/suffixes, I just used genAffixesListOpt. not sure the difference between opt/non-opt, assuming opt is better.

@@ -32,16 +32,17 @@ if __name__ == '__main__':
     parser.add_argument('--now', action='store_true')
     parser.add_argument('--vocab', choices=['small', 'med', 'large', 'full'], default='small')
     parser.add_argument('--vectors', choices=['small', 'med', 'large', 'full'], default='small')
-    parser.add_argument('--lang', choices=['eng', 'tur'], default='eng')
+    parser.add_argument('--lang', choices=['eng', 'tur', 'both'], default='eng')
     parser.add_argument('--supervised', type=float, default=0.0)
+    parser.add_argument('--maxiter', type=int, default=500)
     args = parser.parse_args()
 
     if sys.version_info >= (3, 0):
         raw_input = input  # ghettooooooooo
 
     trainsegmentations, devsegmentations = get_segmentations(args.lang)
-    mc_args = (get_wordvectors(args.vectors, args.lang), get_wordlist(args.vocab, args.lang),
-               *get_prefixes_affixes(args.lang))
+    mc_args = (get_wordvectors(args.vectors, args.lang), get_wordlist(args.vocab, args.lang)) + \
+            get_prefixes_affixes(args.lang)
     mc_kwargs = {'segmentations': trainsegmentations}
 
     if args.command == 'optimize':
@@ -50,21 +51,21 @@ if __name__ == '__main__':
         train = morpho.genTrainingData()
         with open('out_py/dictvectorizer.p', 'wb') as f:
             pickle.dump(morpho.dictvectorizer, f)
-        with open('out_py/dictvectorizer.%s-%s.p' % (args.vocab, args.vectors), 'wb') as f:
+        with open('out_py/dictvectorizer.%s.%s-%s.p' % (args.lang, args.vocab, args.vectors), 'wb') as f:
             pickle.dump(morpho.dictvectorizer, f)
         if args.supervised:
             sups = morpho.genGoldsegTrainingData()
             print('training data saved, optimizing weights')
-            weights = objective.optimize_weights_supervised(*(train + sups), lamb2=args.supervised)
+            weights = objective.optimize_weights_supervised(*(train + sups), lamb2=args.supervised, maxiter=args.maxiter)
             morpho.setWeightVector(weights)
-            with open('out_py/weights.supervised%s.%s-%s.p' %
-                      (args.supervised, args.vocab, args.vectors), 'wb') as f:
+            with open('out_py/weights.%s.supervised%s.%s-%s.p' %
[…]

### 2015-12-09 pysrc/morphochain.py

@@ -39,7 +39,7 @@ class MorphoChain(object):
         self.prefixNeighbours, self.suffixNeighbours = affixNeighbours
         self.weightvector = weightvector
         self.segmentations = segmentations
-        self.translations, self.invTranslations = translations
+        self.translations, self.invTranslations = translations or (None, None)
         self.secondLangChain = secondLangChain
 
     def setWeightVector(self, weightvector):
@@ -92,8 +92,8 @@ class MorphoChain(object):
                     # # FIX: this is currently storing only the last prefix into d['diff']
                     # d['diffpre_' + prefix] = cos_sim
             if affix in self.prefixNeighbours:
-                for n, score in self.prefixNeighbours[affix][:TOPNEIGHBOURS]:
-                    if parent + n in vocab:
+                for n, score in self.prefixNeighbours[affix][:ParentType.TOPNEIGHBOURS]:
+                    if parent + n in self.vocab:
                         d['neighbours_COR_S'] = affix
         else:  # some sort of suffix
             if z.transformtype == ParentType.SUFFIX:
@@ -118,8 +118,8 @@ class MorphoChain(object):
                     # d['diffsuff_' + suffix ] = cos_sim
             # # affix correlation TODO check for each case
             if affix in self.suffixNeighbours:
-                for n, score in self.suffixNeighbours[affix][:TOPNEIGHBOURS]:
-                    if w[:lenparent - len(affix)] + n in vocab:
+                for n, score in self.suffixNeighbours[affix][:ParentType.TOPNEIGHBOURS]:
+                    if w[:lenparent - len(affix)] + n in self.vocab:
                         d['neighbours_COR_S'] = affix
          # parent is not in word list
         if z.parentword not in self.vocab:

### 2015-12-09 pysrc/objective.py

@@ -31,7 +31,7 @@ def output_decorate(func):
     return newf
 
 
-def optimize_weights(X, nzs, widsneighbors, lamb=0, output=True):
+def optimize_weights(X, nzs, widsneighbors, lamb=0, output=True, maxiter=MAX_ITERS):
     '''
     X is a feature matrix, where each row corresponds
     to a (w, z) pair, where the w's appear in order.
@@ -60,12 +60,12 @@ def optimize_weights(X, nzs, widsneighbors, lamb=0, output=True):
         bounds=[(-BOUNDS, BOUNDS)]*X.shape[1],
         approx_grad=False,
         fprime=None,
-        maxiter=MAX_ITERS)[0]
+        maxiter=maxiter)[0]
 
 
 def optimize_weights_supervised(X, nzs, widsneighbors,
                                 Xsup, nzs_sup, cxs_sup,
-                                lamb=0, lamb2=1, output=True):
+                                lamb=0, lamb2=1, output=True, maxiter=MAX_ITERS):
     fus = get_optimizer_fn(X, nzs, widsneighbors, lamb)
     fsup = get_logprob_fn(Xsup, nzs_sup, cxs_sup)
     def fcomb(weights):
@@ -82,7 +82,7 @@ def optimize_weights_supervised(X, nzs, widsneighbors,
         bounds=[(-BOUNDS, BOUNDS)]*X.shape[1],
         approx_grad=False,
         fprime=None,
-        maxiter=MAX_ITERS)[0]
+        maxiter=maxiter)[0]
 
 
 def get_logprob_fn(X, nzs, cxs):

### 2015-12-09 pysrc/suffixes.py

@@ -1,4 +1,4 @@
-import pysrc.fileio
+from pysrc import fileio
 from collections import Counter, defaultdict
 import math
 import numpy as np
@@ -80,39 +80,57 @@ def genAffixesListOpt(wordlist, wordvectors):
         prefixes[prefix] *= (1 + 0.3 * np.linalg.norm(prefixvector[prefix]))
     suff_list = [s[0] for s in suffixes.most_common(100)]
     pref_list = [p[0] for p in prefixes.most_common(100)]
-    print(suff_list)
-    print(pref_list)
+    return suff_list, pref_list
 
 
-def genAffixCorrelation(affixes, wordlist):
+def genAffixCorrelation(affixes, wordlist, fname='../data/prefix_corr3.p', suff=True):
     d = {}
     d2 = {}
     for affix in affixes:
         d[affix] = set()
         affixlen = len(affix)
         for word in wordlist:
-            if word[-affixlen:] == affix:
-                d[affix].add(word[:-affixlen])
+            word_affix = word[-affixlen:] if suff else word[:affixlen]
+            word_root = word[:-affixlen] if suff else word[affixlen:]
+            if word_affix == affix:
+                d[affix].add(word_root)
     for affix in affixes:
         d2[affix] = []
         for affix2 in affixes:
             if affix != affix2:
                 d2[affix].append((affix2, float(len(d[affix] & d[affix2])) / len(d[affix])))
         d2[affix].sort(key = lambda x: x[1])
-    pickle.dump(d2, open("../data/prefix_corr3.p", "wb"))
+    pickle.dump(d2, open(fname, 'wb'))
 
+entr = {'eng': 'en', 'tur': 'tr'}
 
-filename = '../data/wordlist-2010.eng.txt'
+lang  = 'tur'
+size = 'filtered'
 
-wordlist = fileio.read_wordcounts(filename)
[…]

### 2015-12-09 pysrc/__main__.py
small fixes

@@ -88,8 +88,13 @@ if __name__ == '__main__':
 
     if not args.now:
         stdout = sys.stdout
-        with open('out_py/results.%s.%s-%s.txt' % (args.lang, args.vocab, args.vectors), 'wb') as sys.stdout:
+        with open('out_py/results.%s%s.%s-%s.txt' % (args.lang, '.supervised%s' % args.supervised
+                  if args.supervised else '', args.vocab, args.vectors), 'wb') as sys.stdout:
+            print('%s%s %s-%s' % (args.lang, ' supervised(%s)' % args.supervised if args.supervised else '',
+                                  args.vocab, args.vectors))
             morpho.computeAccuracy(devsegmentations)
+        print('%s%s %s-%s' % (args.lang, ' supervised(%s)' % args.supervised if args.supervised else '',
+                              args.vocab, args.vectors))
         sys.stdout = stdout
     morpho.computeAccuracy(devsegmentations)
 

### 2015-12-09 pysrc/__main__.py
touch ups

@@ -93,9 +93,9 @@ if __name__ == '__main__':
             print('%s%s %s-%s' % (args.lang, ' supervised(%s)' % args.supervised if args.supervised else '',
                                   args.vocab, args.vectors))
             morpho.computeAccuracy(devsegmentations)
+        sys.stdout = stdout
         print('%s%s %s-%s' % (args.lang, ' supervised(%s)' % args.supervised if args.supervised else '',
                               args.vocab, args.vectors))
-        sys.stdout = stdout
     morpho.computeAccuracy(devsegmentations)
 
     if args.command == 'run':

### 2015-12-09 pysrc/objective.py

@@ -18,8 +18,11 @@ def output_decorate(func):
         stime = time.time()
         fv, gv = func(weights)
         iteration[0] += 1
-        with open('out_py/weights.p', 'wb') as f:
-            pickle.dump(weights, f)
+        try:
+            with open('out_py/weights.p', 'wb') as f:
+                pickle.dump(weights, f)
+        except:
+            pass
         print('Call %s' % iteration[0])
         print('\tWeights range: %s %s' % (weights.min(), weights.max()))
         print('\tWeights norm (d=%s): %s' % (weights.size, np.linalg.norm(weights)))

### 2015-12-09 pysrc/__main__.py
two language stuff, python 3 compat

@@ -89,13 +89,13 @@ if __name__ == '__main__':
     if not args.now:
         stdout = sys.stdout
         with open('out_py/results.%s%s.%s-%s.txt' % (args.lang, '.supervised%s' % args.supervised
-                  if args.supervised else '', args.vocab, args.vectors), 'wb') as sys.stdout:
+                  if args.supervised else '', args.vocab, args.vectors), 'w') as sys.stdout:
             print('%s%s %s-%s' % (args.lang, ' supervised(%s)' % args.supervised if args.supervised else '',
                                   args.vocab, args.vectors))
             morpho.computeAccuracy(devsegmentations)
+        sys.stdout = stdout
         print('%s%s %s-%s' % (args.lang, ' supervised(%s)' % args.supervised if args.supervised else '',
                               args.vocab, args.vectors))
-        sys.stdout = stdout
     morpho.computeAccuracy(devsegmentations)
 
     if args.command == 'run':

### 2015-12-09 pysrc/__main__.py
store training data

@@ -35,6 +35,7 @@ if __name__ == '__main__':
     parser.add_argument('--lang', choices=['eng', 'tur', 'both'], default='eng')
     parser.add_argument('--supervised', type=float, default=0.0)
     parser.add_argument('--maxiter', type=int, default=500)
+    parser.add_argument('--recalc', action='store_true')
     args = parser.parse_args()
 
     if sys.version_info >= (3, 0):
@@ -48,10 +49,22 @@ if __name__ == '__main__':
     if args.command == 'optimize':
         morpho = MorphoChain(*mc_args, **mc_kwargs)
         print('generating training data')
-        train = morpho.genTrainingData()
+        train_file = 'out_py/%s-train.%s-%s.p' % (args.lang, args.vocab, args.vectors)
+        dictvec_file = 'out_py/dictvectorizer.%s.%s-%s.p' % (args.lang, args.vocab, args.vectors)
+        try:
+            if args.recalc:
+                raise
+            with open(dictvec_file, 'rb') as f:
+                morpho.dictvectorizer = pickle.load(f)
+            with open(train_file, 'rb') as f:
+                train = pickle.load(f)
+        except:
+            train = morpho.genTrainingData()
+            with open(train_file, 'wb') as f:
+                pickle.dump(train, f, pickle.HIGHEST_PROTOCOL)
+            with open(dictvec_file, 'wb') as f:
+                pickle.dump(morpho.dictvectorizer, f)
         with open('out_py/dictvectorizer.p', 'wb') as f:
-            pickle.dump(morpho.dictvectorizer, f)
-        with open('out_py/dictvectorizer.%s.%s-%s.p' % (args.lang, args.vocab, args.vectors), 'wb') as f:
             pickle.dump(morpho.dictvectorizer, f)
         if args.supervised:
             sups = morpho.genGoldsegTrainingData()

### 2015-12-09 pysrc/morphochain.py

@@ -35,7 +35,7 @@ class MorphoChain(object):
         self.dictionary = dictionary
         self.prefixes, self.suffixes = affixes
         self.alphabet = alphabet
-        self.dictvectorizer = dictvectorizer
+        self.dictvectorizer = dictvectorizer or DictVectorizer()
         self.prefixNeighbours, self.suffixNeighbours = affixNeighbours
         self.weightvector = weightvector
         self.segmentations = segmentations

### 2015-12-09 pysrc/__main__.py
results, let me know if you want specific others

@@ -36,6 +36,7 @@ if __name__ == '__main__':
     parser.add_argument('--supervised', type=float, default=0.0)
     parser.add_argument('--maxiter', type=int, default=500)
     parser.add_argument('--recalc', action='store_true')
+    parser.add_argument('--lamb', type=float, default=0.0)
     args = parser.parse_args()
 
     if sys.version_info >= (3, 0):
@@ -48,7 +49,6 @@ if __name__ == '__main__':
 
     if args.command == 'optimize':
         morpho = MorphoChain(*mc_args, **mc_kwargs)
-        print('generating training data')
         train_file = 'out_py/%s-train.%s-%s.p' % (args.lang, args.vocab, args.vectors)
         dictvec_file = 'out_py/dictvectorizer.%s.%s-%s.p' % (args.lang, args.vocab, args.vectors)
         try:
@@ -59,6 +59,7 @@ if __name__ == '__main__':
             with open(train_file, 'rb') as f:
                 train = pickle.load(f)
         except:
+            print('generating training data')
             train = morpho.genTrainingData()
             with open(train_file, 'wb') as f:
                 pickle.dump(train, f, pickle.HIGHEST_PROTOCOL)
@@ -68,15 +69,15 @@ if __name__ == '__main__':
             pickle.dump(morpho.dictvectorizer, f)
         if args.supervised:
             sups = morpho.genGoldsegTrainingData()
-            print('training data saved, optimizing weights')
-            weights = objective.optimize_weights_supervised(*(train + sups), lamb2=args.supervised, maxiter=args.maxiter)
+            print('training data acquired, optimizing weights')
+            weights = objective.optimize_weights_supervised(*(train + sups), lamb=args.lamb, lamb2=args.supervised, maxiter=args.maxiter)
             morpho.setWeightVector(weights)
             with open('out_py/weights.%s.supervised%s.%s-%s.p' %
                       (args.lang, args.supervised, args.vocab, args.vectors), 'wb') as f:
                 pickle.dump(weights, f)
         else:
-            print('training data saved, optimizing supervised weights')
-            weights = objective.optimize_weights(*train, maxiter=args.maxiter)
+            print('training data acquired, optimizing supervised weights')
+            weights = objective.optimize_weights(*train, lamb=args.lamb, maxiter=args.maxiter)
             morpho.setWeightVector(weights)
             with open('out_py/weights.%s.%s-%s.p' % (args.lang, args.vocab, args.vectors), 'wb') as f:
                 pickle.dump(weights, f)

### 2015-12-09 pysrc/morphochain.py

@@ -35,7 +35,7 @@ class MorphoChain(object):
         self.dictionary = dictionary
         self.prefixes, self.suffixes = affixes
         self.alphabet = alphabet
-        self.dictvectorizer = dictvectorizer or DictVectorizer()
+        self.dictvectorizer = dictvectorizer
         self.prefixNeighbours, self.suffixNeighbours = affixNeighbours
         self.weightvector = weightvector
         self.segmentations = segmentations

### 2015-12-09 pysrc/duolingo.py
kimchi fried rice

@@ -2,6 +2,7 @@ from morphochain import MorphoChain, ParentTransformation, ParentType
 import string
 import editdistance
 
+
 class TwoLangMorphoChain(MorphoChain):
     def __init__(self, wordvectors, vocab, affixes, affixNeighbours, translations,
                  dictionary=None,
@@ -58,5 +59,14 @@ class TwoLangMorphoChain(MorphoChain):
 
         return list(canddict.values())
 
-
-
+    def getFeatures(self, w, z, maxCosSimilarity=None):
+        d = super(TwoLangMorphoChain, self).getFeatures(w, z, maxCosSimilarity)
+        if z.olangconfidence is not None:
+            d['olangconfidence'] = z.olangconfidence
+        else:
+            d['no_translated_parents'] = 1
+        if z.transformtype != ParentType.OLANG and z.olangconfidence is not None:
+            d['repeat'] = z.olangconfidence
+        else:
+            d['no_repeat'] = 1
+        return d

### 2015-12-09 pysrc/morphochain.py

@@ -77,7 +77,10 @@ class MorphoChain(object):
         parent = z.parentword
         lenparent = len(z.parentword)
         # affix
-        if z.transformtype == ParentType.PREFIX:
+        if z.transformtype == ParentType.OLANG:
+            pass  # features for OLANG will be in duolingo, this is just to prevent
+                  # base features from messing up
+        elif z.transformtype == ParentType.PREFIX:
             affix = w[:-lenparent]
             # list of prefixes  #TODO: Maxlength of suffix is a param??
             if len(affix) > MAX_PREF_LEN or affix not in self.prefixes:
@@ -167,8 +170,9 @@ class MorphoChain(object):
     def genNeighbors(self, w, k=5):
         k = min(k, (len(w)) // 2)
         ne = set([w])
+
         def swap(word, i):
-            return word[:i] + word[i+1] + word[i] + word[i+2:]
+            return word[:i] + word[i + 1] + word[i] + word[i + 2:]
         for i in range(k):
             ne.add(swap(w, i))
         for i in range(len(w) - k - 1, len(w) - 1):

### 2015-12-09 pysrc/duolingo.py
eggs

@@ -1,7 +1,14 @@
 from morphochain import MorphoChain, ParentTransformation, ParentType
 import string
 import editdistance
+from pysrc import fileio
+import pickle
 
+
+TWOLANG_LANG = 'tur'
+TWOLANG_VEC = 'full'
+TWOLANG_VOCAB = 'med'
+TWOLANG_SUPERVISED = 0.5
 
 class TwoLangMorphoChain(MorphoChain):
     def __init__(self, wordvectors, vocab, affixes, affixNeighbours, translations,
@@ -16,35 +23,6 @@ class TwoLangMorphoChain(MorphoChain):
         self.translations, self.invTranslations = translations
         self.secondLangChain = secondLangChain
         self.init2LangCache()
-
-    def parentHeuristic(word, parent):
-        dist = editdistance.eval(word, parent)
-        return dist < 6 and dist < len(word) / 2 and dist < len(parent) / 2
-
-    def init2LangCache(self):
-        tWordToEngParents = {}
-        for tWord in self.invTranslations:  # every Turkish word that can be translated to
-            turkishParents = [(z.parentword, c) for z, c in self.secondLangChain.predict(tWord)]
-            englishParents = {}
-            turkishParents.sort(key=lambda x: x[1])  # sort by increasing confidence
-            for tParent, c in turkishParents:
-                if tParent in self.invTranslations:
-                    for eParent in self.invTranslations[tParent]:
-                        englishParents[eParent] = c  # englishParents[eParent] is maximum confidence
-
-            tWordToEngParents[tWord] = sorted(englishParents.items(), key=lambda x: x[1])
-
-        self.eWordToEngParents = {}
-        for eWord, tTranslations in self.translations.items():
-            possibleParents = []
-            for tWord in tTranslations:
-                possibleParents.extend(tWordToEngParents[tWord])
-            possibleParents.sort(key=lambda x: x[1])
-            englishParents = {}
-            for parent, score in possibleParents:
-                englishParents[parent] = score
-            self.eWordToEngParents[eWord] = [(parent, score) for parent, score in possibleParents.items()
-                                             if self.parentHeuristic(eWord, parent)]
 
     def genCandidates(self, word):
         ocandidates = super(TwoLangMorphoChain, self).genCandidates(word)
[…]

### 2015-12-09 paper/paper.tex

@@
 % TODO explain existing code
+\section{Introduction}
+Morphologically related words\dots %%TODO
 
@@
 
-\section{Using \LaTeX with PDF Figures}
-
-This is a sample document for use with pdflatex, which is
-a program that is included with the Miktex distribution
-that directly produces PDF files from \LaTeX sources.
-To run \LaTeX on this file, you need the following files:
-\begin{enumerate}
-\item templatePDF.tex (this file)
-\item figure.pdf (the figure file)
-\item simpleConference.sty (style file)
-\item refs.bib (bibiliography file)
-\end{enumerate}
-\noindent
-To create a PDF file, execute the following commands:
-\begin{enumerate}
-\item pdflatex templatePDF
-\item bibtex templatePDF
-\item pdflatex templatePDF
-\item pdflatex templatePDF
-\end{enumerate}
-\noindent
-Yes (strangely) it is necessary to run pdflatex three times.
-The result will be a PDF file (plus several other files that \LaTeX
-produces).  You will need a mechanism, of course, for executing
-commands on the command line. If you are using Windows, I recommend
-installing Cygwin and using its bash shell.
-
-\section{How to Include Vergil Diagrams as Figures}
-
-Suppose you wish to include a figure, like that in figure \ref{fig-label}.
-The simplest mechanism is to install Adobe Acrobat, which includes
-a ``printer'' called ``Acrobat Distiller.'' Printing to this printer
-creates a PDF file, which can be included in a document as shown
-here.  To include Ptolemy II models \cite{PtolemyVol1:04},
-just print to the distiller from within Vergil and reference
-the PDF file in your \LaTeX document.
-
-There is a bit more work to do, however.
-The file that is produced by the distiller represents
-a complete page, not the individual figure.
-You can open it in using Acrobat (version 5.0 or later),
-and select Document $\rightarrow$ Crop Pages from the menu.
-In the resulting dialog, check ``Remove White Margins.''
-Save the modified PDF file in a file and then reference
-it in the \LaTeX file as shown in this example.
-
-An alternative is to generate EPS (encapsulated postscript),
-but the process is much more complex and fragile.
-I recommend using pdflatex and Adobe Acrobat.
-
-\bibliographystyle{abbrv}
-\bibliography{refs}
 \end{document}

### 2015-12-09 pysrc/duolingo.py
Stuff

@@ -4,7 +4,7 @@ import editdistance
 
 
 class TwoLangMorphoChain(MorphoChain):
-    def __init__(self, wordvectors, vocab, affixes, affixNeighbours, translations,
+    def __init__(self, wordvectors, vocab, affixes, affixNeighbours, eWordToEngParents,
                  dictionary=None,
                  alphabet=string.ascii_lowercase, dictvectorizer=None,
                  weightvector=None, segmentations=None,
@@ -12,10 +12,9 @@ class TwoLangMorphoChain(MorphoChain):
         super(TwoLangMorphoChain, self).__init__(
             wordvectors, vocab, affixes, affixNeighbours, dictionary,
             alphabet, dictvectorizer, weightvector,
-            segmentations, translations)
-        self.translations, self.invTranslations = translations
+            segmentations)
         self.secondLangChain = secondLangChain
-        self.init2LangCache()
+        self.eWordToEngParents = eWordToEngParents
 
     def parentHeuristic(word, parent):
         dist = editdistance.eval(word, parent)

### 2015-12-09 pysrc/morphochain.py

@@ -28,8 +28,7 @@ class ParentType():
 class MorphoChain(object):
     def __init__(self, wordvectors, vocab, affixes, affixNeighbours, dictionary=None,
                  alphabet=string.ascii_lowercase, dictvectorizer=None,
-                 weightvector=None, segmentations=None,
-                 secondLangChain=None):
+                 weightvector=None, segmentations=None):
         self.wordvectors = wordvectors
         self.vocab = vocab
         self.dictionary = dictionary
@@ -39,7 +38,6 @@ class MorphoChain(object):
         self.prefixNeighbours, self.suffixNeighbours = affixNeighbours
         self.weightvector = weightvector
         self.segmentations = segmentations
-        self.secondLangChain = secondLangChain
 
     def setWeightVector(self, weightvector):
         self.weightvector = weightvector

### 2015-12-09 pysrc/duolingo.py
gold suffixes

@@ -1,7 +1,7 @@
+from pysrc import *
 from morphochain import MorphoChain, ParentTransformation, ParentType
 import string
 import editdistance
-from pysrc import fileio
 import pickle
 
 
@@ -83,8 +83,8 @@ def init2LangCache(translations, invTranslations, secondLangChain):
 
 
 if __name__ == "__main__":
-    mc_args = (fileio.get_wordvectors(TWOLANG_VEC, TWOLANG_LANG), fileio.get_wordlist(TWOLANG_VOCAB, TWOLANG_LANG)) + \
-            fileio.get_prefixes_affixes(TWOLANG_LANG)
+    mc_args = (get_wordvectors(TWOLANG_VEC, TWOLANG_LANG), get_wordlist(TWOLANG_VOCAB, TWOLANG_LANG)) + \
+            get_prefixes_affixes(TWOLANG_LANG)
     mc_kwargs = {}
     with open('out_py/dictvectorizer.%s.%s-%s.p' % (TWOLANG_LANG, TWOLANG_LANG, TWOLANG_LANG), 'rb') as f:
         mc_kwargs['dictvectorizer'] = pickle.load(f)
@@ -93,7 +93,7 @@ if __name__ == "__main__":
         mc_kwargs['weightvector'] = pickle.load(f)
 
     turkishMorpho = MorphoChain(*mc_args, **mc_kwargs)
-    translations, invTranslations = fileio.read_dictionary('sozluk.txt')
+    translations, invTranslations = read_dictionary('sozluk.txt')
     cache = init2LangCache(translations, invTranslations, turkishMorpho)
     with open('data/eWordToEParents.p', 'wb') as f:
         pickle.dump(cache, f)

### 2015-12-09 pysrc/morphochain.py
gold suffixes

@@ -130,7 +130,7 @@ class MorphoChain(object):
         else:
             d['parent_in_word_list'] = math.log10(self.vocab[z.parentword])
         # presence in English dictionary
-        # d['parent_in_dict'] = z.parentword in self.dictionary
+        #d['parent_in_dict'] = z.parentword in self.dictionary
 
         #TODO USE dp -- extend C(w) using existing affixes and word2vec
         return d

### 2015-12-09 pysrc/__main__.py
some neatening of __main__

@@ -1,4 +1,5 @@
 from pysrc import *
+import pysrc.duolingo as duolingo
 from pysrc.morphochain import MorphoChain
 import rlcompleter
 import readline
@@ -49,17 +50,22 @@ if __name__ == '__main__':
     mc_kwargs = {'segmentations': trainsegmentations}
 
     MCCls = MorphoChain
+    filesuffix = '.%s.%s-%s' % (args.lang, args.vocab, args.vectors)
+    ofilesuffix = filesuffix if not args.supervised else \
+        '.%s.supervised%s.%s-%s' % (args.lang, args.supervised, args.vocab, args.vectors)
+
     if args.twolang:
         MCCls = duolingo.TwoLangMorphoChain
         with open('out_py/eWordToEParents.p') as f:
             eWordToEparents = pickle.load(f)
             mc_args.append(eWordToEParents)
-
+        filesuffix = '.twolang' + filesuffix
+        ofilesuffix = '.twolang' + ofilesuffix
 
     if args.command == 'optimize':
         morpho = MCCls(*mc_args, **mc_kwargs)
         train_file = 'out_py/%s-train.%s-%s.p' % (args.lang, args.vocab, args.vectors)
-        dictvec_file = 'out_py/dictvectorizer.%s.%s-%s.p' % (args.lang, args.vocab, args.vectors)
+        dictvec_file = 'out_py/dictvectorizer%s.p' % filesuffix
         try:
             if args.recalc:
                 raise
@@ -81,15 +87,12 @@ if __name__ == '__main__':
             print('training data acquired, optimizing weights')
             weights = objective.optimize_weights_supervised(*(train + sups), lamb=args.lamb, lamb2=args.supervised, maxiter=args.maxiter)
             morpho.setWeightVector(weights)
-            with open('out_py/weights.%s.supervised%s.%s-%s.p' %
-                      (args.lang, args.supervised, args.vocab, args.vectors), 'wb') as f:
-                pickle.dump(weights, f)
         else:
             print('training data acquired, optimizing supervised weights')
             weights = objective.optimize_weights(*train, lamb=args.lamb, maxiter=args.maxiter)
             morpho.setWeightVector(weights)
-            with open('out_py/weights.%s.%s-%s.p' % (args.lang, args.vocab, args.vectors), 'wb') as f:
-                pickle.dump(weights, f)
+        with open('out_py/weights%s.p' % ofilesuffix, 'wb') as f:
+            pickle.dump(weights, f)
 
[…]

### 2015-12-09 pysrc/duolingo.py

@@ -1,7 +1,7 @@
-from morphochain import MorphoChain, ParentTransformation, ParentType
+from pysrc.morphochain import MorphoChain, ParentTransformation, ParentType
+from pysrc import fileio
 import string
 import editdistance
-from pysrc import fileio
 import pickle
 
 

### 2015-12-09 pysrc/duolingo.py
oops

@@ -1,5 +1,5 @@
 from pysrc import *
-from morphochain import MorphoChain, ParentTransformation, ParentType
+from pysrc.morphochain import MorphoChain, ParentTransformation, ParentType
 import string
 import editdistance
 import pickle

### 2015-12-09 pysrc/__main__.py
oops fix scoring, wow such f1

@@ -126,9 +126,9 @@ if __name__ == '__main__':
         word = raw_input("Enter word: ")
         while True:
             if word == 'ACCURACY_TRAIN':
-                morpho.computeAccuracy()
+                morpho.computeAccuracy(verbose=True)
             elif word =='ACCURACY_DEVEL':
-                morpho.computeAccuracy(devsegmentations)
+                morpho.computeAccuracy(devsegmentations, True)
             elif word == 'RELOAD' and args.now:
                 loadweights()
             elif word == 'EXIT':

### 2015-12-10 pysrc/duolingo.py
update 2lang

@@ -8,7 +8,7 @@ import pickle
 TWOLANG_LANG = 'tur'
 TWOLANG_VEC = 'full'
 TWOLANG_VOCAB = 'med'
-TWOLANG_SUPERVISED = 0.5
+TWOLANG_SUPERVISED = 20.0
 
 class TwoLangMorphoChain(MorphoChain):
     def __init__(self, wordvectors, vocab, affixes, affixNeighbours, eWordToEngParents,
@@ -51,7 +51,10 @@ class TwoLangMorphoChain(MorphoChain):
 
 def parentHeuristic(word, parent):
     dist = editdistance.eval(word, parent)
-    return dist < 6 and dist < len(word) / 2 and dist < len(parent) / 2
+    lendiff = len(word) - len(parent)
+    return lendiff >= -1 and\
+            dist < 6 and\
+            dist <= (len(word) + 2) / 2
 
 
 def init2LangCache(translations, invTranslations, secondLangChain):
@@ -76,7 +79,7 @@ def init2LangCache(translations, invTranslations, secondLangChain):
         englishParents = {}
         for parent, score in possibleParents:
             englishParents[parent] = score
-        eWordToEngParents[eWord] = [(parent, score) for parent, score in possibleParents.items()
+        eWordToEngParents[eWord] = [(parent, score) for parent, score in englishParents.items()
                                             if parentHeuristic(eWord, parent)]
     return eWordToEngParents
 
@@ -85,14 +88,14 @@ if __name__ == "__main__":
     mc_args = (get_wordvectors(TWOLANG_VEC, TWOLANG_LANG), get_wordlist(TWOLANG_VOCAB, TWOLANG_LANG)) + \
             get_prefixes_affixes(TWOLANG_LANG)
     mc_kwargs = {}
-    with open('out_py/dictvectorizer.%s.%s-%s.p' % (TWOLANG_LANG, TWOLANG_LANG, TWOLANG_LANG), 'rb') as f:
+    with open('out_py/dictvectorizer.%s.%s-%s.p' % (TWOLANG_LANG, TWOLANG_VOCAB, TWOLANG_VEC), 'rb') as f:
         mc_kwargs['dictvectorizer'] = pickle.load(f)
     with open('out_py/weights.%s%s.%s-%s.p' % (TWOLANG_LANG, '.supervised%s' % TWOLANG_SUPERVISED
[…]

### 2015-12-10 pysrc/__main__.py
two lang stuff

@@ -56,7 +56,7 @@ if __name__ == '__main__':
 
     if args.twolang:
         MCCls = duolingo.TwoLangMorphoChain
-        with open('out_py/eWordToEParents.p') as f:
+        with open('data/eWordToEParents.p') as f:
             eWordToEparents = pickle.load(f)
             mc_args.append(eWordToEParents)
         filesuffix = '.twolang' + filesuffix

### 2015-12-10 pysrc/duolingo.py

@@ -39,7 +39,10 @@ class TwoLangMorphoChain(MorphoChain):
     def getFeatures(self, w, z, maxCosSimilarity=None):
         d = super(TwoLangMorphoChain, self).getFeatures(w, z, maxCosSimilarity)
         if z.olangconfidence is not None:
-            d['olangconfidence'] = z.olangconfidence
+            if z.transformtype == ParentType.STOP:
+                d['stop_olangconfidence'] = z.olangconfidence
+            else:
+                d['olangconfidence'] = z.olangconfidence
         else:
             d['no_translated_parents'] = 1
         if z.transformtype != ParentType.OLANG and z.olangconfidence is not None:

### 2015-12-10 paper/paper.tex
started paper model

@@
 Morphologically related words\dots %%TODO
+
+\section{Background Work}
+Our model builds on work done by Narasimhan et al.
+(2015) in constructing \emph{morphological chains} to model the morphological segmentation of words in a language in an unsupervised manner.
+The idea behind morphological chains is that complex words are constructed by attaching morphemes simpler words.
+Thus, complex words have words from which they are directly derived, which we call \emph{parent} words.
+Likewise, the derived word is a \emph{child} word.
+A morphological chain is a sequence of words such that consecutive pairs form a parent-child relationship.
+For example, the word \emph{unsustainable} can be constructed from \emph{sustain}, as demonstrated in the chain \emph{sustain} $\rightarrow$ \emph{sustainable} $\rightarrow$ \emph{unsustainable}.
+Words without parents are called \emph{base} words.
+In our example, \emph{sustain} is a base word.
+
+Since a parent word can have multiple children (e.g., \emph{play} $\rightarrow$ \emph{plays} and \emph{play} $\rightarrow$ \emph{played}), words can be part of multiple chains.
+Thus multiple chains can share segments.
+Narasimhan makes use of this shared information by constructing a model which analyzes parent-child pairs.
+
+\subsection{Model}
+
+A log-linear model is used to evaluate different pairs.
+For this, a feature mapping $\phi: \mathcal W \times \mathcal Z \rightarrow \mathbb R^d$ to compute a corresponding weight vector $\theta\in\mathbb R^d$.
+$\mathcal W$ is the set of words being trained on, and $\mathcal Z$ is the set of \emph{candidates} for the parents the words in $\mathcal W$.
+For a word $w\in\mathcal W$, $\mathcal Z$ is constructed by splitting $w$ at many points.
+To capture orthographic changes in the parent word as it undergoes the derivation from $z\in Z$ to $w$ (eg., \emph{believe} $\rightarrow$ \emph{believing}), the type of transformation is kept as part of the candidate.
+Thus the candidates take the form (\emph{parent}, \emph{type}), where the type is the type of transformation.
+
+Note that parent words usually undergo changes only when attaching suffixes.
+Thus there is one \emph{Prefix} class, but there are a variety of suffix classes.
+When acquiring a suffix, the parent word can: (1) undergo no change (\emph{class} $\rightarrow$ \emph{classes}), (2) repeat a character (\emph{star} $\rightarrow$ \emph{starring}), (3) delete a character (\emph{believe} $\rightarrow$ \emph{believing}), (4) modify a character (\emph{parry} $\rightarrow$ \emph{parried}).
+These correspond to the candidates (\emph{class}, \emph{Suffix}), (\emph{star}, \emph{Repeat}), (\emph{believe}, \emph{Delete}), and (\emph{parry}, \emph{Modify}).
+Finally, there is a \emph{Stop} type, which is used to signify that the word is a base word.
+
+\subsection{Features}
+
+The model uses a variety of features covering orthographic and semantic aspects of the word-candidate pairs. These features are computed for pairs $(w,z) \in \mathcal W \times \mathcal Z$.
 

### 2015-12-10 pysrc/accuracy.py
better distance shit

@@ -3,7 +3,7 @@ import pprint
 
 
 def seg_points(segs):
-    return set(np.cumsum([len(i) for i in segs])[:-1])
+    return set(np.cumsum([len(i) for i in segs if i != '~'])[:-1])
 
 
 def score(gold_segs_list, predictions_segs_list, verbose=False):

### 2015-12-10 pysrc/duolingo.py

@@ -1,7 +1,6 @@
 from pysrc import *
 from pysrc.morphochain import MorphoChain, ParentTransformation, ParentType
 import string
-import editdistance
 import pickle
 
 
@@ -52,11 +51,27 @@ class TwoLangMorphoChain(MorphoChain):
         return d
 
 
+def distance(word, parent):
+    lencommonprefix = 0
+    lencommonsuffix = 0
+    for a, b in zip(word, parent):
+        if a != b:
+            break
+        lencommonprefix += 1
+
+    for a, b in zip(word[lencommonprefix:][::-1], parent[lencommonprefix:][::-1]):
+        if a != b:
+            break
+        lencommonsuffix += 1
+
+    return max(len(word), len(parent)) - lencommonprefix - lencommonsuffix
+
+
 def parentHeuristic(word, parent):
-    dist = editdistance.eval(word, parent)
+    dist = distance(word, parent)
     lendiff = len(word) - len(parent)
     return lendiff >= -1 and\
-            dist < 6 and\
+            dist < 5 and\
             dist <= (len(word) + 2) / 2
 
 

### 2015-12-10 paper/paper.tex

@@
-\documentclass[10pt,twocolumn]{article} 
+\documentclass[11pt,twocolumn]{article} 
 \usepackage{simpleConference}
@@
 
+Note that by constructing morphological chains rather than just finding segmentations, more information is gained about the words tested.
+Generating the chain, in addition to segmenting a word, also finds the base word and predicts the order in which the morphemes attach to the base word.
+
 \subsection{Model}
+
+In the unsupervised version of our model, we observe words in a wordlist with their word count of occurrences.
+Additionally, we have access to a large corpus of text from which we can obtain semantic information.
+Our objective is to generate morphological segmentations, which we do by generating morphological chains, as given by Narasimhan et al.
 
@@
 The model uses a variety of features covering orthographic and semantic aspects of the word-candidate pairs. These features are computed for pairs $(w,z) \in \mathcal W \times \mathcal Z$.
+
+\begin{description}
+    \item[Affixes] Prefixes and suffixes appear in many words.
+For a given word with its potential parent, the affix removed can be compared against a precompiled list of potential affixes.
+This allows the model to learn which affixes are real and apply them to other cases with the same potential affix.
+    \item[Affix Correlation] Similar to comparing affixes across words, we can use the joint distribution of words and their potential affixes.
+There is a correlation between affixes that can usually attach to the same stem, often corresponding to the part of speech.
+For example, the participles formed from the suffixes -ing and -ed usually occur together.
+For candidates with an affix that is correlated with another affix, we can check if the parent with the other affix also occurs in the observed wordlist.
+    \item[Semantic Similarity] Morphologically related words should have similar meanings.
+We measure the \emph{cosine similarity} between the word and its candidate parent.
+% TODO wordvector citation?
+        It has been shown that words can be mapped to word vectors in a reasonably-sized dimensional space such that semantically similar 
+\end{description}
 

### 2015-12-10 paper/paper.tex
paper

--- 2015-12-10 paper/paper.tex
@@
 
+\section{Experimental Validation}
+
+\subsection{Data}
+
+We used the datasets provided by 2010 Morpho Challenges. For English, this includes
+a wordlist with word frequencies of approximately 878,000 words scraped from the
+Wortschatz collection from the University of Leipzig, CLEF, and the Europarl corpus
+as well as gold standard segmentation training and development sets of approximately 
+1000 randomly selected words. Gold standard segmentation is based on the CELEX database.
+For Turkish, the word list contains approximately
+
 \end{document}

### 2015-12-10 paper/paper.tex

--- 2015-12-10 paper/paper.tex
@@
 For a word $w\in\mathcal W$, $\mathcal Z$ is constructed by splitting $w$ at many points.
-To capture orthographic changes in the parent word as it undergoes the derivation from $z\in Z$ to $w$ (eg., \emph{believe} $\rightarrow$ \emph{believing}), the type of transformation is kept as part of the candidate.
-Thus the candidates take the form (\emph{parent}, \emph{type}), where the type is the type of transformation.
+To capture orthographic changes in the parent word as it undergoes the derivation from $z\in Z$ to $w$ (eg., \emph{believe} $\rightarrow$ \emph{believing}), the type of transition is kept as part of the candidate.
+Thus the candidates take the form (\emph{parent}, \emph{type}), where the type is the type of transition.
 
@@
 Finally, there is a \emph{Stop} type, which is used to signify that the word is a base word.
+
+\subsection{Word Vectors}
+
+% TODO wordvector citation?
+It has been shown that words can be mapped to word vectors in a reasonably-sized dimensional space such that semantically similar words align in similar directions.
+Thus we can compute a \emph{cosine similarity} as a distance metric between words which captures semantic similarity.
+Given vector representations of two words, the cosine similarity is measured as the dot product between the word vectors of those words.
+
+In our work, we generate 200-dimensional vectors for words by using Word2Vec, which looks at word alignment  and co-occurrence patterns across many documents.
+In generating our vectors, we use text from Wikipedia along with a corpus of news articles.
 
@@
     \item[Semantic Similarity] Morphologically related words should have similar meanings.
-We measure the \emph{cosine similarity} between the word and its candidate parent.
-% TODO wordvector citation?
-        It has been shown that words can be mapped to word vectors in a reasonably-sized dimensional space such that semantically similar 
+We measure the cosine similarity between the word and its candidate parent as a feature to represent semantic similarity.
+    \item[Transformations] As discussed above, some derivations involve a change to the parent word.
+To allow for non-concatenative morphology, we use binary features which capture the type of transformation.
+We use the same types of transformations for features as we do for generating candidates: repetitions, deletions, and modifications.
+The set of binary features are the cartesian product of the type of
 \end{description}

### 2015-12-10 paper/paper.tex

@@
 
-\section{Experimental Validation}
-
-\subsection{Data}
-
-We used the datasets provided by 2010 Morpho Challenges. For English, this includes
-a wordlist with word frequencies of approximately 878,000 words scraped from the
-Wortschatz collection from the University of Leipzig, CLEF, and the Europarl corpus
-as well as gold standard segmentation training and development sets of approximately 
-1000 randomly selected words. Gold standard segmentation is based on the CELEX database.
-For Turkish, the word list contains approximately
-
 \end{document}

### 2015-12-10 paper/paper.tex
features fin

@@
 We use the same types of transformations for features as we do for generating candidates: repetitions, deletions, and modifications.
-The set of binary features are the cartesian product of the type of
+The set of binary features are the cartesian product of the type of transformation with the characters transformed.
+    \item[Wordlist] Most of the time, we want the parents to be valid words.
+To this effect, we compare the candidate parent against the wordlist, and add a feature for the log of the word count.
+Additionally, we set a binary feature corresponding to whether the word was found at all in the wordlist.
+    \item[Stop Features] We have a number of miscellaneous features targeted at finding the end of chains.
+These include orthographic information like the length of the parent, unigrams and bigrams at the beginning and end of the parent, and the highest cosine similarity between the word and any of its candidate parents.
 \end{description}
@@
         L_S(\theta; D) &= \log{\prod_{w, z \in D} P(z|w)} \\
-                     &= \sum_{w, z \in D} \left(\theta \cdot \phi(w, z) - \log\left(\sum_{z' \in C(w)} e^{\theta \cdot \phi(w, z')} \right)\right)
+                       &= \sum_{w, z \in D} \Bigg(\theta \cdot \phi(w, z) \Bigg. \\
+                       &\qquad\Bigg. - \log\left(\sum_{z' \in C(w)} e^{\theta \cdot \phi(w, z')} \right)\Bigg)
    \end{split}
@@
     \begin{split}
-        \nabla_\theta L_S &= \sum_{w, z \in D} \left(\phi(w, z) - \frac{\sum_{z' \in C(w)} \phi(w, z') e^{\theta \cdot \phi(w, z')} }{\sum_{z' \in C(w)} e^{\theta \cdot \phi(w, z')}}\right)
+        \nabla_\theta L_S &= \sum_{w, z \in D} \Bigg(\phi(w, z) \Bigg. \\
+                          &\qquad\left. - \frac{\sum_{z' \in C(w)} \phi(w, z') e^{\theta \cdot \phi(w, z')} }{\sum_{z' \in C(w)} e^{\theta \cdot \phi(w, z')}}\right)
     \end{split}

### 2015-12-10 paper/paper.tex
graphs

--- 2015-12-10 paper/paper.tex
@@
 
+\section{Experimental Validation}
+
+\subsection{Data}
+
+We used the datasets provided by 2010 Morpho Challenges. For English, this includes
+a wordlist with word frequencies of approximately 878,000 words scraped from the
+Wortschatz collection from the University of Leipzig, CLEF, and the Europarl corpus
+as well as gold standard segmentation training and development sets of approximately 
+1000 randomly selected words. Gold standard segmentation is based on the CELEX database.
+For Turkish, the word list contains approximately
+
+\begin{center}
+    \begin{tabular}{ | l | l | l | l | l |}
+    \hline
+    Language & Method & Precision & Recall & F1 \\ \hline
+      & Narasimhan & 0.807 & 0.722 & 0.762 \\ 
+    & Unsupervised & -  & - & - \\ 
+    English & Semi-supervised & - & - & - \\ 
+     & Two language unsupervised & - & - & - \\ 
+     & Two language semi-supervised & - & - & - \\ \hline
+     & Narasimhan & 0.743 & 0.520 & 0.612 \\ 
+    & Unsupervised & - & - & - \\ 
+    Turkish & Semi-supervised & - & - & -  \\ 
+     & Two language unsupervised  & - & - & -  \\ 
+     & Two language semi-supervised  & - & - & -  \\ \hline
+    \end{tabular}
+\end{center}
+
+\begin{center}
+    \begin{tabular}{ | l | l | l | l | l | l |}
+    \hline
+    Language & \multicolumn{2}{|c|}{\textbf{Correct Segmentations}} & \multicolumn{3}{|c|}{\textbf{Incorrect Segmentations}}\\ \hline
+      & Narasimhan & 0.807 & 0.722 & 0.762& - \\ 
+    & Unsupervised & -  & - & - & -\\ 
+    English & Semi-supervised & - & - & - & -\\ 
+     & Two language unsupervised & - & - & -& - \\ 
+     & Two language semi-supervised & - & - & - & -\\ \hline
+     & Narasimhan & 0.743 & 0.520 & 0.612 & -\\ 
+    & Unsupervised & - & - & - & -\\ 
+    Turkish & Semi-supervised & - & - & -  & -\\ 
+     & Two language unsupervised  & - & - & -  & -\\ 
+     & Two language semi-supervised  & - & - & -  & -\\ \hline
+    \end{tabular}
+\end{center}
+
 \end{document}

### 2015-12-10 paper/paper.tex

--- 2015-12-10 paper/paper.tex
@@
 
+\begin{center}
+    \begin{tabular}{ | l | l | l | l | l |}
+    \hline
+    Language & Method & Precision & Recall & F1 \\ \hline
+      & Narasimhan & 0.807 & 0.722 & 0.762 \\ 
+    & Unsupervised & -  & - & - \\ 
+    English & Semi-supervised & - & - & - \\ 
+     & Two language unsupervised & - & - & - \\ 
+     & Two language semi-supervised & - & - & - \\ \hline
+     & Narasimhan & 0.743 & 0.520 & 0.612 \\ 
+    & Unsupervised & - & - & - \\ 
+    Turkish & Semi-supervised & - & - & -  \\ 
+     & Two language unsupervised  & - & - & -  \\ 
+     & Two language semi-supervised  & - & - & -  \\ \hline
+    \end{tabular}
+\end{center}
+
+\begin{center}
+    \begin{tabular}{ | l | l | l | l | l | l |}
+    \hline
+    Language & \multicolumn{2}{|c|}{\textbf{Correct Segmentations}} & \multicolumn{3}{|c|}{\textbf{Incorrect Segmentations}}\\ \hline
+      & Narasimhan & 0.807 & 0.722 & 0.762& - \\ 
+    & Unsupervised & -  & - & - & -\\ 
+    English & Semi-supervised & - & - & - & -\\ 
+     & Two language unsupervised & - & - & -& - \\ 
+     & Two language semi-supervised & - & - & - & -\\ \hline
+     & Narasimhan & 0.743 & 0.520 & 0.612 & -\\ 
+    & Unsupervised & - & - & - & -\\ 
+    Turkish & Semi-supervised & - & - & -  & -\\ 
+     & Two language unsupervised  & - & - & -  & -\\ 
+     & Two language semi-supervised  & - & - & -  & -\\ \hline
+    \end{tabular}
+\end{center}
+
 \end{document}

### 2015-12-10 paper/paper.tex
paper stuf

--- 2015-12-10 paper/paper.tex
@@
 Thus there is one \emph{Prefix} class, but there are a variety of suffix classes.
-When acquiring a suffix, the parent word can: (1) undergo no change (\emph{class} $\rightarrow$ \emph{classes}), (2) repeat a character (\emph{star} $\rightarrow$ \emph{starring}), (3) delete a character (\emph{believe} $\rightarrow$ \emph{believing}), (4) modify a character (\emph{parry} $\rightarrow$ \emph{parried}).
+When acquiring a suffix, the parent word can:
+\begin{enumerate}
+    \item undergo no change (\emph{class} $\rightarrow$ \emph{classes})
+    \item repeat a character (\emph{star} $\rightarrow$ \emph{starring})
+    \item delete a character (\emph{believe} $\rightarrow$ \emph{believing})
+    \item modify a character (\emph{parry} $\rightarrow$ \emph{parried}).
+\end{enumerate}
 These correspond to the candidates (\emph{class}, \emph{Suffix}), (\emph{star}, \emph{Repeat}), (\emph{believe}, \emph{Delete}), and (\emph{parry}, \emph{Modify}).
@@
 In our work, we generate 200-dimensional vectors for words by using Word2Vec, which looks at word alignment  and co-occurrence patterns across many documents.
-In generating our vectors, we use text from Wikipedia along with a corpus of news articles.
+In generating our vectors, we use text from Wikipedia,
+to generate the english word vectors, along with the BOUN corpus of news articles to generate Turkish word vectors.
 
@@
 
-\section{Experimental Validation}
-
-\subsection{Data}
-
-We used the datasets provided by 2010 Morpho Challenges. For English, this includes
-a wordlist with word frequencies of approximately 878,000 words scraped from the
-Wortschatz collection from the University of Leipzig, CLEF, and the Europarl corpus
-as well as gold standard segmentation training and development sets of approximately 
-1000 randomly selected words. Gold standard segmentation is based on the CELEX database.
-For Turkish, the word list contains approximately
-
-\begin{center}
-    \begin{tabular}{ | l | l | l | l | l |}
-    \hline
-    Language & Method & Precision & Recall & F1 \\ \hline
-      & Narasimhan & 0.807 & 0.722 & 0.762 \\ 
-    & Unsupervised & -  & - & - \\ 
-    English & Semi-supervised & - & - & - \\ 
-     & Two language unsupervised & - & - & - \\ 
-     & Two language semi-supervised & - & - & - \\ \hline
-     & Narasimhan & 0.743 & 0.520 & 0.612 \\ 
-    & Unsupervised & - & - & - \\ 
-    Turkish & Semi-supervised & - & - & -  \\ 
-     & Two language unsupervised  & - & - & -  \\ 
-     & Two language semi-supervised  & - & - & -  \\ \hline
-    \end{tabular}
-\end{center}
-
-\begin{center}
-    \begin{tabular}{ | l | l | l | l | l | l |}
-    \hline
-    Language & \multicolumn{2}{|c|}{\textbf{Correct Segmentations}} & \multicolumn{3}{|c|}{\textbf{Incorrect Segmentations}}\\ \hline
-      & Narasimhan & 0.807 & 0.722 & 0.762& - \\ 
-    & Unsupervised & -  & - & - & -\\ 
-    English & Semi-supervised & - & - & - & -\\ 
-     & Two language unsupervised & - & - & -& - \\ 
-     & Two language semi-supervised & - & - & - & -\\ \hline
-     & Narasimhan & 0.743 & 0.520 & 0.612 & -\\ 
-    & Unsupervised & - & - & - & -\\ 
-    Turkish & Semi-supervised & - & - & -  & -\\ 
-     & Two language unsupervised  & - & - & -  & -\\ 
-     & Two language semi-supervised  & - & - & -  & -\\ \hline
-    \end{tabular}
-\end{center}
-
 \end{document}

### 2015-12-10 pysrc/duolingo.py

@@ -39,15 +39,16 @@ class TwoLangMorphoChain(MorphoChain):
         d = super(TwoLangMorphoChain, self).getFeatures(w, z, maxCosSimilarity)
         if z.olangconfidence is not None:
             if z.transformtype == ParentType.STOP:
-                d['stop_olangconfidence'] = z.olangconfidence
+                d['stop_olangconfidence'] = z.olangconfidence * 0.01
             else:
-                d['olangconfidence'] = z.olangconfidence
+                d['olangconfidence'] = z.olangconfidence * 0.01
         else:
             d['no_translated_parents'] = 1
         if z.transformtype != ParentType.OLANG and z.olangconfidence is not None:
-            d['repeat'] = z.olangconfidence
+            d['repeat'] = z.olangconfidence * 0.01
         else:
             d['no_repeat'] = 1
+        d['lendiff%s' % (len(w) - len(z))] = 1
         return d
 
 
@@ -78,7 +79,7 @@ def parentHeuristic(word, parent):
 def init2LangCache(translations, invTranslations, secondLangChain):
     tWordToEngParents = {}
     for tWord in invTranslations:  # every Turkish word that can be translated to
-        turkishParents = [(z.parentword, c) for z, c in secondLangChain.predict(tWord)]
+        turkishParents = [(z.parentword, c) for z, c in secondLangChain.predict_logprobs(tWord)]
         englishParents = {}
         turkishParents.sort(key=lambda x: x[1])  # sort by increasing confidence
         for tParent, c in turkishParents:

### 2015-12-10 pysrc/accuracy.py
fix accuracy

@@ -19,8 +19,8 @@ def score(gold_segs_list, predictions_segs_list, verbose=False):
             continue
         predpoints = len(pred_segs) - 1
         predTotal += predpoints
-        bestc = 0
-        bestg = 0
+        bestc = -1
+        bestg = -1
         for i, gold_segs in enumerate(gold_segs_list[word][0]):
             gset = seg_points(gold_segs)
             pset = seg_points(pred_segs)
@@ -31,7 +31,7 @@ def score(gold_segs_list, predictions_segs_list, verbose=False):
                 bestc = num_correct_segs
                 bestg = num_gold_segs
         if bestc == bestg == predpoints:
-            correctSegs.append(pred_seg)
+            correctSegs.append((pred_seg, '/'.join(gold_segs)))
         else:
             incorrectSegs.append((pred_seg, "/".join(gold_segs)))
 

### 2015-12-10 paper/paper.tex
paper

--- 2015-12-10 paper/paper.tex
@@
 1000 randomly selected words. Gold standard segmentation is based on the CELEX database.
-For Turkish, the word list contains approximately
+For Turkish, the word list contains approximately 670,000 words, with gold standards
+from a morphological parser developed at Boğaziçi University. The Turkish wordlist
+was obtained from a significantly smaller corpus -- at 1 million words, as compared
+to an English corpus of 18 million.
+
+We tested on 4 different sizes 
+of wordvector 
+small is 2.5k
+med is 
+large is 160k
+full is 878k
+
+\subsection{Evaluation Model}
 
@@
     Language & \multicolumn{2}{|c|}{\textbf{Correct Segmentations}} & \multicolumn{3}{|c|}{\textbf{Incorrect Segmentations}}\\ \hline
-      & Narasimhan & 0.807 & 0.722 & 0.762& - \\ 
-    & Unsupervised & -  & - & - & -\\ 
-    English & Semi-supervised & - & - & - & -\\ 
-     & Two language unsupervised & - & - & -& - \\ 
+     & Word & Segmentation & Word & Predicted & Correct\\ \hline
+    & 'suburbanite' & 'suburb/an/ite' & 'provokingly' & 'provok/ing/ly' & 'provoking/ly'  \\ 
+    'buffeted' & 'buffeted' & -  & - & - & -\\ 
+    "ewers'" &  "ewer/s/'" & - & - & - & -\\ 
+    'yowling' & 'yowl/ing' & - & - & -& - \\ 
      & Two language semi-supervised & - & - & - & -\\ \hline

### 2015-12-10 paper/paper.tex
learning stuff

--- 2015-12-10 paper/paper.tex
@@
 \begin{document} 
-
 \title{Extending Morphological Chains with Supervised Learning and Cross-Language Features}
@@
 
-\section{Background Work}
+\section{Unsupervised Model}
 Our model builds on work done by Narasimhan et al.
@@
 Thus there is one \emph{Prefix} class, but there are a variety of suffix classes.
-When acquiring a suffix, the parent word can:
-\begin{enumerate}
-    \item undergo no change (\emph{class} $\rightarrow$ \emph{classes})
-    \item repeat a character (\emph{star} $\rightarrow$ \emph{starring})
-    \item delete a character (\emph{believe} $\rightarrow$ \emph{believing})
-    \item modify a character (\emph{parry} $\rightarrow$ \emph{parried}).
-\end{enumerate}
+When acquiring a suffix, the parent word can: (1) undergo no change (\emph{class} $\rightarrow$ \emph{classes}), (2) repeat a character (\emph{star} $\rightarrow$ \emph{starring}), (3) delete a character (\emph{believe} $\rightarrow$ \emph{believing}), (4) modify a character (\emph{parry} $\rightarrow$ \emph{parried}).
 These correspond to the candidates (\emph{class}, \emph{Suffix}), (\emph{star}, \emph{Repeat}), (\emph{believe}, \emph{Delete}), and (\emph{parry}, \emph{Modify}).
@@
 In our work, we generate 200-dimensional vectors for words by using Word2Vec, which looks at word alignment  and co-occurrence patterns across many documents.
-In generating our vectors, we use text from Wikipedia,
-to generate the english word vectors, along with the BOUN corpus of news articles to generate Turkish word vectors.
+In generating our vectors, we use text from Wikipedia along with a corpus of news articles.
 
@@
 \end{description}
+
+\subsection{Unsupervised Learning}
+Recall that in our model, we want to find weights $\theta$ for the feature vectors $\phi(w,z)$, where $w\in\mathcal W$ and $z\in\mathcal Z$.
+Define the probabilities of a word-candidate pair $(w,z)$ as $P(w,z)\propto e^{\theta\cdot\phi(w,z)}$.
+Then the probability of a candidate $z$ ocuring given $w$ is
+\begin{equation}
+    P(z|w) = \frac{e^{\theta\cdot\phi(w,z)}} {\sum_{z'\in C(w)} e^{\theta\cdot\phi(w,z')}}
+\end{equation}
+
+Let our set of observed words be $D$.
+We approach this by trying to maximize the likelihood of observing the words in $D$ from the space of all constructible strings from the alphabet, $\Sigma^*$.
+We use the EM-algorithm to maximize the likelihood over $\theta$ as given by
+
+\begin{equation}
+    \begin{split}
+        L(\theta; D) &= \prod_{w^*\in D} P(w^*) \\
+        &= \prod_{w^*\in D} \sum_{z\in C(w^*)} P(w^*, z) \\
+        &= \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in\Sigma^*}\sum_{z\in C(w) e^{\theta\cdot\phi(w,z)}}} \\
+    \end{split}
+\end{equation}
+
+We cannot compute $L(\theta; D)$ directly since we cannot compute over all strings in $\Sigma^*$.
+Instead, we approximate this distribution by considering only strings which are similar to those encountered in $D$.
+
+We use the method of Contrastive Estimation (Smith and Eisner, 2005) and substitute the space of all strings $\Sigma^*$ with neighbors of each word, $N(w)$.
+Toward this end, we transpose pairs of consecutive letters of $w$ near both ends of the word.
+We also do both simultaneously.
+Together these form our set of neighbors for $w$.
+The neighbors form a proxy for $\Sigma^*$, and represent the set of strings we want to reduce the probability of seeing in our model because they are not observed.
+This has the benefit of providing contrast in the structure of words while not requiring the model to look at the entire $\Sigma^*$.
+
+With this substitution, we can formulate the contrastive likelihood as
+\begin{equation}
+    L_C(\theta; D) = \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}}
+\end{equation}
 
@@
 there are many cases in which the potential parent is not actually generated;
-for example, the morphological parent of "feet" is the singular "foot";
+for example, the morphological parent of ``feet'' is the singular ``foot'';
 however, there is no prefix or suffix that turns one into the other, and
-therefore the parent for "feet" is not generated by our system.
+therefore the parent for ``feet'' is not generated by our system.
 
@@
 for the same words, for example in Turkish the corresponding translations
-for "feet" and "foot" are "ayaklar" and "ayak". A properly trained Turkish
-morphological parent model will correctly detect "ayak" as a parent of "ayaklar",
+for ``feet'' and ``foot'' are ``ayaklar'' and ``ayak''. A properly trained Turkish
+morphological parent model will correctly detect ``ayak'' as a parent of ``ayaklar'',
 and can be used to suggest morphological parents in English after translation.

### 2015-12-10 paper/paper.tex
two lang and stuff

--- 2015-12-10 paper/paper.tex
@@
 from a morphological parser developed at Boğaziçi University. The Turkish wordlist
-was obtained from a significantly smaller corpus -- at 1 million words, as compared
-to an English corpus of 18 million.
+was obtained from a significantly smaller corpus --- at 1 million words, compared
+to the English corpus of 18 million.
 
-We tested on 4 different sizes 
+We tested on 4 different sizes of wordvectors:
 of wordvector 
@@
 
+\begin{table*}
     \begin{center}
@@
         \end{tabular}
+        \caption{Results 'n shit TODO}
     \end{center}
+\end{table*}
 
+\begin{table*}
     \begin{center}
@@
                      & Word & Segmentation & Word & Predicted & Correct\\ \hline
-    & 'suburbanite' & 'suburb/an/ite' & 'provokingly' & 'provok/ing/ly' & 'provoking/ly'  \\ 
-    'buffeted' & 'buffeted' & -  & - & - & -\\ 
-    "ewers'" &  "ewer/s/'" & - & - & - & -\\ 
-    'yowling' & 'yowl/ing' & - & - & -& - \\ 
+            English & suburbanite & suburb/an/ite & provokingly & provok/ing/ly & provoking/ly  \\ 
+                    & buffeted & buffeted & -  & - & -\\ 
+                    & ewers' &  ewer/s/' & - & - & -\\ 
+                    & yowling & yowl/ing & - & -& - \\ 
                     & Two language semi-supervised & - & - & - & -\\ \hline
@@
         \end{tabular}
+        \caption{Results 'n shit TODO}
     \end{center}
+\end{table*}
 

### 2015-12-10 pysrc/morphochain.py

@@ -1,4 +1,5 @@
 from collections import namedtuple, Counter
+import sys
 import math
 import string
 import itertools
@@ -47,10 +48,13 @@ class MorphoChain(object):
         Return dict from possible parents to feature dict
         """
         parentsAndFeatures = {}
+        parentt = None
         for z in self.genCandidates(w):
+            if z.transformtype == ParentType.STOP:
+                parentt = z
+                continue
             parentsAndFeatures[z] = self.getFeatures(w, z)
-        z = ParentTransformation(w, ParentType.STOP, None)
-        parentsAndFeatures[z] = self.getFeatures(
+        parentsAndFeatures[parentt] = self.getFeatures(
             w, z, max([d['cos'] for d in parentsAndFeatures.values()] or [0]))
         return parentsAndFeatures
 
@@ -161,6 +165,7 @@ class MorphoChain(object):
         for x in range(1, (len(word) + 2) // 2):
             parent = word[x:]
             candidates.append(ParentTransformation(parent, ParentType.PREFIX, None))
+        candidates.append(ParentTransformation(word, ParentType.STOP, None))
         # Stopping condition handled in getParentsAndFeatures
         # candidates.append(ParentTransformation(word, ParentType.STOP))
         return candidates
@@ -293,7 +298,7 @@ class MorphoChain(object):
         elif candidate[1] == ParentType.PREFIX:
             return word[:-p_len] + SEG_SEP + self.genSeg(parent)
         else:
-            raise
+            raise Exception((word, parent))
 
     def similarity(self, w1, w2):
         if w1 not in self.wordvectors or w2 not in self.wordvectors:

### 2015-12-10 paper/paper.tex
paper

--- 2015-12-10 paper/paper.tex
@@
 \section{Introduction}
-Morphologically related words\dots %%TODO
+The division of a word into morphemes--basic units of meaning--- encodes information about
+something
+When performing morphological analysis, there are a few different cases to consider: words that share similar
+meaning but appear in diffrent syntactic contexts (``German" ---``Germanic"), new words generated from
+the combination of two separate words (``bookcase"|``book"---``case"), words with dissimilar meaning but
+derived from one another ``insuffcient"---``sufficient"). In the first and last case, words are often related
+by the addition of preffixes, suffixes, and other affixes. This allows us to form morphological chains - which
+link a base word to its derivative through a series of affixations.
+We wish to
+
+%%TODO
 
@@
 
-explanation blah
+We want to maximize the probability of the document $D$, where $w$ is a word in $D$ and $c$ is
+its segmentation. As above, we 
 %TODO better explanation
@@
 from a morphological parser developed at Boğaziçi University. The Turkish wordlist
-was obtained from a significantly smaller corpus --- at 1 million words, compared
-to the English corpus of 18 million.
+was obtained from a significantly smaller corpus -- at 1 million words, as compared
+to an English corpus of 18 million.
 
-We tested on 4 different sizes of wordvectors:
+We tested on 4 different sizes 
 of wordvector 
@@
 
-\begin{table*}
 \begin{center}
@@
     \end{tabular}
-        \caption{Results 'n shit TODO}
 \end{center}
-\end{table*}
 
-\begin{table*}
 \begin{center}
@@
      & Word & Segmentation & Word & Predicted & Correct\\ \hline
-            English & suburbanite & suburb/an/ite & provokingly & provok/ing/ly & provoking/ly  \\ 
-                    & buffeted & buffeted & -  & - & -\\ 
-                    & ewers' &  ewer/s/' & - & - & -\\ 
-                    & yowling & yowl/ing & - & -& - \\ 
+    & 'suburbanite' & 'suburb/an/ite' & 'provokingly' & 'provok/ing/ly' & 'provoking/ly'  \\ 
+    'buffeted' & 'buffeted' & -  & - & - & -\\ 
+    "ewers'" &  "ewer/s/'" & - & - & - & -\\ 
+    'yowling' & 'yowl/ing' & - & - & -& - \\ 
      & Two language semi-supervised & - & - & - & -\\ \hline
@@
     \end{tabular}
-        \caption{Results 'n shit TODO}
 \end{center}
-\end{table*}
 

### 2015-12-10 paper/paper.tex
log likelihood equations

--- 2015-12-10 paper/paper.tex
@@
 
+With a regularization term, this becomes
+\begin{equation}
+    \begin{split}
+        L_C(\theta; D) &= \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}} \\
+        &\qquad - \lambda \|\theta\|^2
+    \end{split}
+\end{equation}
+
+Our EM-algorithm will then operate on the log-likelihood, which takes the following form:
+\begin{equation}
+    \begin{split}
+        L_L(\theta; D) &= \sum_{w^*\in D} \Bigg(\log \sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)} \Bigg.\\
+        &\qquad\Bigg. -\log \sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)} \Bigg) \\
+        &\qquad - \lambda \|\theta\|^2
+    \end{split}
+\end{equation}
+
+This has the gradient given by
+\begin{equation}
+    \begin{split}
+        &\frac{\partial L_L(\theta; D)}{\partial \theta_j} \\
+        &= \sum_{w^*\in D} \Bigg( \frac{\sum_{z\in C(w^*)} \phi_j(w^*,z)\cdot e^{\theta\cdot\phi(w^*,z)}} {\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} \Bigg. \\
+        &\qquad -\frac{\sum_{w\in N(w^*)} \sum_{z\in C(w)} \phi_j(w,z)\cdot e^{\theta\cdot\phi(w,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}} \Bigg) \\
+        &\qquad -2\lambda\theta_j
+    \end{split}
+\end{equation}
+
+We optimize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm (Byrd et al., 1995).
+
 \section{Semi-supervised Learning with Known Word Segmentations}

### 2015-12-10 paper/paper.tex
paper bib stuff

--- 2015-12-10 paper/paper.tex
@@
 \usepackage{simpleConference}
+\usepackage{natbib}
 \usepackage{amsmath}
@@
 \section{Unsupervised Model}
-Our model builds on work done by Narasimhan et al.
-(2015) in constructing \emph{morphological chains} to model the morphological segmentation of words in a language in an unsupervised manner.
+Our model builds on work done by \citet{Narasimhan15} in constructing \emph{morphological chains} to model the morphological segmentation of words in a language in an unsupervised manner.
 The idea behind morphological chains is that complex words are constructed by attaching morphemes simpler words.
@@
 
-Therefore, the likelihood of the training data is given as such:
+We want to maximize the likelihood of the document $D$, where $w$ is a word in $D$ and $c$ is
+its segmentation:
 \begin{equation}
-    \prod_{w, c \in D} P(c | w) = \prod_{w, c \in D}\left(\prod_{w', z' \in MC(w)} P(z | w)\right)
+    \prod_{w, c \in D} P(c | w) = \prod_{w, c \in D}\left(\prod_{w', z' \in MC(w, c)} P(z' | w')\right)
 \end{equation}
+where $MC(w)$ represents the parents and transitions in the morphological chain of $w$ going up to the root,
+constructed using its segmentation.
 
-We want to maximize the probability of the document $D$, where $w$ is a word in $D$ and $c$ is
-its segmentation. As above, we 
-%TODO better explanation
-
-Therefore, the log-likelihood of our training data can be expressed as
+We can simplify this product by letting $TR(D)$ be all pairs of words and parents implied in each of
+the morphological chains of all words in $D$ ---
+then, the log-likelihood of our training data can be expressed as
 \begin{equation}
     \begin{split}
-        L_S(\theta; D) &= \log{\prod_{w, z \in D} P(z|w)} \\
-                       &= \sum_{w, z \in D} \Bigg(\theta \cdot \phi(w, z) \Bigg. \\
+        L_S(\theta; D) &= \log{\prod_{w, z \in TR(D)} P(z|w)} \\
+                       &= \sum_{w, z \in TR(D)} \Bigg(\theta \cdot \phi(w, z) \Bigg. \\
                        &\qquad\Bigg. - \log\left(\sum_{z' \in C(w)} e^{\theta \cdot \phi(w, z')} \right)\Bigg)
@@
 \end{equation}
-where $w, z$ represent every morphological chain transition implied in the training data.
 
@@
 
+\bibliography{paper}
+\bibliographystyle{plainnat}
+
 \end{document}

### 2015-12-10 paper/paper.tex
more citation stuff

@@
 Thus multiple chains can share segments.
-Narasimhan makes use of this shared information by constructing a model which analyzes parent-child pairs.
+\citeauthor{Narasimhan15} makes use of this shared information by constructing a model which analyzes parent-child pairs.
 
@@
 Additionally, we have access to a large corpus of text from which we can obtain semantic information.
-Our objective is to generate morphological segmentations, which we do by generating morphological chains, as given by Narasimhan et al.
+Our objective is to generate morphological segmentations, which we do by generating morphological chains, as given by \citeauthor{Narasimhan15}
 
@@
 Additionally, we extend our feature set to include whether or not each candidate was
-generated by a turkish translation (a strong indicator for parent-ness) and how confident
+generated by a Turkish translation (a strong indicator for parent-ness) and how confident
 the Turkish model was in selecting a parent.
@@
             Language & Method & Precision & Recall & F1 \\ \hline
-                     & Narasimhan & 0.807 & 0.722 & 0.762 \\ 
+                     & \citeauthor{Narasimhan15} & 0.807 & 0.722 & 0.762 \\ 
                      & Unsupervised & -  & - & - \\ 
@@
                     & Two language semi-supervised & - & - & - \\ \hline
-                    & Narasimhan & 0.743 & 0.520 & 0.612 \\ 
+                    & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 \\ 
                     & Unsupervised & - & - & - \\ 
@@
                     & yowling & yowl/ing & - & -& - \\ 
-                    & Two language semi-supervised & - & - & - & -\\ \hline
-                    & Narasimhan & 0.743 & 0.520 & 0.612 & -\\ 
+                    & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 & -\\ 
                     & Unsupervised & - & - & - & -\\ 

### 2015-12-10 paper/paper.tex
paper

--- 2015-12-10 paper/paper.tex
@@
                      & Word & Segmentation & Word & Predicted & Correct\\ \hline
-    & 'suburbanite' & 'suburb/an/ite' & 'provokingly' & 'provok/ing/ly' & 'provoking/ly'  \\ 
-    'buffeted' & 'buffeted' & -  & - & - & -\\ 
-    "ewers'" &  "ewer/s/'" & - & - & - & -\\ 
-    'yowling' & 'yowl/ing' & - & - & -& - \\ 
+            English & suburbanite & suburb/an/ite & provokingly & provok/ing/ly & provoking/ly  \\ 
+                    & buffeted & buffeted & -  & - & -\\ 
+                    & ewers' &  ewer/s/' & - & - & -\\ 
+                    & yowling & yowl/ing & - & -& - \\ 
                     & Two language semi-supervised & - & - & - & -\\ \hline
                     & Narasimhan & 0.743 & 0.520 & 0.612 & -\\ 
-    & Unsupervised & - & - & - & -\\ 
-    Turkish & Semi-supervised & - & - & -  & -\\ 
-     & Two language unsupervised  & - & - & -  & -\\ 
-     & Two language semi-supervised  & - & - & -  & -\\ \hline
+            Turkish & Unsupervised & - & - & - & -\\ 
+                    & Semi-supervised & - & - & -  & -\\ 
+                    & Two language unsupervised  & - & - & -  & -\\ \hline
+                    & - & - & - & -  & -\\ 
+                    & - & - & - & -  & -\\ 
+    English (dual-language) & - & - & - & -  & -\\ 
+                    & - & - & - & -  & -\\ 
+                    & - & - & - & -  & -\\ \hline
                 \end{tabular}
+        \caption{Results 'n shit TODO}
     \end{center}
+\end{table*}
 

### 2015-12-10 paper/paper.tex
OOPS

--- 2015-12-10 paper/paper.tex
@@
                     & yowling & yowl/ing & - & -& - \\ 
-                    & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 & -\\ 
-                    & Unsupervised & - & - & - & -\\ 
-            Turkish & Semi-supervised & - & - & -  & -\\ 
-                    & Two language unsupervised  & - & - & -  & -\\ 
                     & Two language semi-supervised & - & - & - & -\\ \hline
+                    & Narasimhan & 0.743 & 0.520 & 0.612 & -\\ 
+            Turkish & Unsupervised & - & - & - & -\\ 
+                    & Semi-supervised & - & - & -  & -\\ 
+                    & Two language unsupervised  & - & - & -  & -\\ \hline
+                    & - & - & - & -  & -\\ 
+                    & - & - & - & -  & -\\ 
+    English (dual-language) & - & - & - & -  & -\\ 
+                    & - & - & - & -  & -\\ 
+                    & - & - & - & -  & -\\ \hline
                 \end{tabular}

### 2015-12-10 paper/paper.tex
stuf

--- 2015-12-10 paper/paper.tex
@@
 In generating our vectors, we use text from Wikipedia,
-to generate the english word vectors, along with the BOUN corpus of news articles to generate Turkish word vectors.
+to generate the English word vectors, along with the BOUN corpus of news articles to generate Turkish word vectors.
 
@@
 where $T_{ET}$ returns the set of possible Turkish translations
-for an english word; $MP_T$ returns the set of possible parents for some Turkish word,
-$T_{TE}$ returns the set of possible english translations for an english word,
+for an English word; $MP_T$ returns the set of possible parents for some Turkish word,
+$T_{TE}$ returns the set of possible English translations for an English word,
 and this potentially large set of candidates is pruned with heuristics judging
@@
                      & \citeauthor{Narasimhan15} & 0.807 & 0.722 & 0.762 \\ 
-                     & Unsupervised & -  & - & - \\ 
+                     & Unsupervised (Baseline) & -  & - & - \\ 
             English & Semi-supervised & - & - & - \\ 
@@
                     & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 \\ 
-                    & Unsupervised & - & - & - \\ 
+                    & Unsupervised (Baseline) & - & - & - \\ 
             Turkish & Semi-supervised & - & - & -  \\ 
@@
         \end{tabular}
-        \caption{Results 'n shit TODO}
+        \caption{Accuracies of Various Models}
     \end{center}
@@
             English & suburbanite & suburb/an/ite & provokingly & provok/ing/ly & provoking/ly  \\ 
-                    & buffeted & buffeted & -  & - & -\\ 
-                    & ewers' &  ewer/s/' & - & - & -\\ 
-                    & yowling & yowl/ing & - & -& - \\ 
-                    & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 & -\\ 
-                    & Unsupervised & - & - & - & -\\ 
-            Turkish & Semi-supervised & - & - & -  & -\\ 
-                    & Two language unsupervised  & - & - & -  & -\\ 
-                    & Two language semi-supervised  & - & - & -  & -\\ \hline
+                    & buffeted & buffet/ed & invalidated  & in/validat/ed & in/valid/at/ed \\ 
+                    & ewers' &  ewer/s/' & dutifully & dutiful/ly & duti/ful/ly\\ 
+                    & yowling & yowl/ing & ignominiously & ignominious/ly& ignomni/ous/ly \\ 
+            \hline
+                    & - & - & - & - & -\\ 
+            Turkish & - & - & - & - & -\\ 
+                    & - & - & - & - & -\\ 
+                    & - & - & - & - & -\\ \hline
         \end{tabular}
-        \caption{Results 'n shit TODO}
+        \caption{Examples of Correct and Incorrect Segmentations}
     \end{center}
 \end{table*}
+
+\section{Results}
+We found that utilizing information from the gold segmentations, we were able to substantially increase our
+performance, from an accuracy of 65.5\% to 81.7\% with English, and from 47.9\% to 70.3\% with Turkish.
+
+We found that adding the additional Turkish candidates and features were actually detrimental to the 
+unsupervised case
 

### 2015-12-10 paper/paper.tex
more results

--- 2015-12-10 paper/paper.tex
@@
 
+Because we have chains where different chains can share segments, it is natural to consider the graph formulation where each word is represented by a node and each parent-child transition is represented by a directed edge.
+Then if we consider only the most likely parent for a word,\footnote{Words are not restricted to having a single valid parent (or segmentation).
+In cases where there are multiple parents (segmentations), we care only that the algorithm selects one of them.} the resulting graph forms a forest, and our objective is to find the path to from a given node to the root in an unsupervised manner without any explicit information about any of the edges.
+
 Note that by constructing morphological chains rather than just finding segmentations, more information is gained about the words tested.
@@
 \begin{enumerate}
-    \item undergo no change (\emph{class} $\rightarrow$ \emph{classes})
+    \item undergo no change (\emph{bike} $\rightarrow$ \emph{bikes})
     \item repeat a character (\emph{star} $\rightarrow$ \emph{starring})
@@
 \end{enumerate}
-These correspond to the candidates (\emph{class}, \emph{Suffix}), (\emph{star}, \emph{Repeat}), (\emph{believe}, \emph{Delete}), and (\emph{parry}, \emph{Modify}).
+These correspond to the candidates (\emph{bike}, \emph{Suffix}), (\emph{star}, \emph{Repeat}), (\emph{believe}, \emph{Delete}), and (\emph{parry}, \emph{Modify}).
 Finally, there is a \emph{Stop} type, which is used to signify that the word is a base word.
@@
 \begin{equation}
+    \label{equ:prob_par}
     P(z|w) = \frac{e^{\theta\cdot\phi(w,z)}} {\sum_{z'\in C(w)} e^{\theta\cdot\phi(w,z')}}
@@
 We approach this by trying to maximize the likelihood of observing the words in $D$ from the space of all constructible strings from the alphabet, $\Sigma^*$.
-We use the EM-algorithm to maximize the likelihood over $\theta$ as given by
+We use the EM-algorithm to iteratively maximize the likelihood over $\theta$ as given by
 
@@
 
-We optimize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm (Byrd et al., 1995).
+We maximize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm (Byrd et al., 1995).
+
+\subsection{Prediction}
+The EM-algorithm yields an optimal $\theta^*$ of weights of the features generated from $D$.
+Then given a test word $w$, we can predict the probability of a given parent candidate $z$ by computing $P(z|w)=e^{\theta^*\cdot\phi(w,z)}$ as in \autoref{equ:prob_par}.
+To predict the parent of $w$, we pick the MLE candidate from parent candidates generated for $w$, which we can generate in the same way as we did while training.
+
+To generate a morphological chain for $w$, we recursively predict the MLE parent for $w$, then the MLE parent of that prediction, and so forth until the \emph{Stop} candidate is predicted.
+Together, this sequence forms our chain.
+A morphological segmentation can be constructed from the chain by splitting $w$ at every point corresponding to where each of the edges in the chain splits its child word.
 
 \section{Semi-supervised Learning with Known Word Segmentations}
+In a semi-supervised learning model, we still have access to a large wordlist of observed words as well as the corpus for semantic information.
+In addition, we have a small amount of pre-segmented words (about 2\% of the size of the wordlist).
+We use these pre-segmented words to generate
+
 In addition to maximizing the log-likelihood of the known wordlist,
@@
                     & Narasimhan & 0.743 & 0.520 & 0.612 & -\\ 
-            Turkish & Unsupervised & - & - & - & -\\ 
-                    & Semi-supervised & - & - & -  & -\\ 
-                    & Two language unsupervised  & - & - & -  & -\\ \hline
-                    & - & - & - & -  & -\\ 
-                    & - & - & - & -  & -\\ 
-    English (dual-language) & - & - & - & -  & -\\ 
-                    & - & - & - & -  & -\\ 
-                    & - & - & - & -  & -\\ \hline
+                    & Unsupervised & - & - & - & -\\ 
+            Turkish & Semi-supervised & - & - & -  & -\\ 
+                    & Two language unsupervised  & - & - & -  & -\\ 
+                    & Two language semi-supervised  & - & - & -  & -\\ \hline
         \end{tabular}

### 2015-12-10 paper/paper.tex

--- 2015-12-10 paper/paper.tex
@@
                     & - & - & - & - & -\\ \hline
+                    & - & - & - & -  & -\\ 
+                    & - & - & - & -  & -\\ 
+    English (dual-language) & - & - & - & -  & -\\ 
+                    & - & - & - & -  & -\\ 
+                    & - & - & - & -  & -\\ \hline
                 \end{tabular}

### 2015-12-10 paper/paper.tex
stuff

--- 2015-12-10 paper/paper.tex
@@
 
-We use the method of Contrastive Estimation (Smith and Eisner, 2005) and substitute the space of all strings $\Sigma^*$ with neighbors of each word, $N(w)$.
+We use the method of Contrastive Estimation \citep{Smith05} and substitute the space of all strings $\Sigma^*$ with neighbors of each word, $N(w)$.
 Toward this end, we transpose pairs of consecutive letters of $w$ near both ends of the word.
@@
 
-We optimize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm (Byrd et al., 1995).
+We optimize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm.
 
@@
 
-Additionally, we extend our feature set to include whether or not each candidate was
+Additionally, we extend our feature set with features regarding the translation link between
+the word and potential parent, to include information such as whether or not each candidate was
 generated by a Turkish translation (a strong indicator for parent-ness) and how confident
 the Turkish model was in selecting a parent.
+
+Therefore, our model becomes:
+\begin{equation}
+    \begin{split}
+        P(w) &= \sum_{z \in C'(w)} P(w,z) \\
+             &= \frac{\sum_{z \in C'(w)} e^{\theta \cdot \phi(w, z, T(w), T(z))}}{\sum_{w' \in N(w)} \sum_{z \in C'(w')} e^{\theta \cdot \phi(w', z', T(w'), T(z'))}}
+    \end{split}
+\end{equation}
+where $C'(w)$ is the potential parents of $w$ updated with potential irregular translations,
+and $T(w)$ is the set of Turkish translations of $w$.
 
@@
 
-We found that adding the additional Turkish candidates and features were actually detrimental to the 
-unsupervised case
-
+We found that adding the additional Turkish candidates and features regarding translation
+did not help accuracy of training very much; this is likely due to the incompleteness of the translation
+dictionary only being able to provide potential translations for very few words, and due to the ambiguity
+of translation, it's very hard to tell whether or not something would make a good parent of a word after
+translating the word into Turkish, taking the parent in Turkish, and translating it back.
 \bibliography{paper}

### 2015-12-10 paper/paper.tex
start of gold segmentations

--- 2015-12-10 paper/paper.tex
@@
 \usepackage{amssymb}
+\usepackage{enumitem}
 \usepackage{url,hyperref}
@@
 When acquiring a suffix, the parent word can:
-\begin{enumerate}
+\begin{enumerate}[label=(\arabic*)]
     \item undergo no change (\emph{bike} $\rightarrow$ \emph{bikes})
@@
 \end{enumerate}
+
 These correspond to the candidates (\emph{bike}, \emph{Suffix}), (\emph{star}, \emph{Repeat}), (\emph{believe}, \emph{Delete}), and (\emph{parry}, \emph{Modify}).
@@
 Thus we can compute a \emph{cosine similarity} as a distance metric between words which captures semantic similarity.
-Given vector representations of two words, the cosine similarity is measured as the dot product between the word vectors of those words.
+Given vector representations of two words, the cosine similarity is measured as the ratio of the dot product between the word vectors and the product of the norms of the word vectors.
 
@@
 Then the probability of a candidate $z$ ocuring given $w$ is
+
 \begin{equation}
@@
 With this substitution, we can formulate the contrastive likelihood as
+
 \begin{equation}
@@
 With a regularization term, this becomes
+
 \begin{equation}
@@
 Our EM-algorithm will then operate on the log-likelihood, which takes the following form:
+
 \begin{equation}
@@
 This has the gradient given by
+
 \begin{equation}
@@
 In addition, we have a small amount of pre-segmented words (about 2\% of the size of the wordlist).
-We use these pre-segmented words to generate
+We use these pre-segmented words to generate probable morphological chains, which we can then incorporate into the likelihood model.
+Specifically, we want to maximize the joint distribution of observing the words in the wordlist along with observing the segmentations of the words from the list of pre-segmented words, with some weighting on each of them.
 
+\subsection{Generating Chains from Segmentations}
+Based on the correct segmentation of a word, we construct a probable chain.
+Assuming no infixes, each parent must be formed by removing segments at either the beginning or the end of the word.
+In generating morphological chains from segmentations, we assume most segmentations are derived from single affix transitions.
+Our method cannot generate the correct chain for compound words where both parts are derived from parent words, for example.
+
+Given a segmentation, we have two candidates for the segmentation of the parent word:
+\begin{enumerate}[label=(\arabic*)]
+    \item
+\end{enumerate}
+
+\subsection{Semi-supervised Model}
 In addition to maximizing the log-likelihood of the known wordlist,
@@
 its segmentation:
+
 \begin{equation}
@@
 then, the log-likelihood of our training data can be expressed as
+
 \begin{equation}
@@
 We can also express its gradient as such:
+
 \begin{equation}
@@
 along with an L2 regularization term, we get our objective:
+
 \begin{equation}
@@
 we generate a set of possible parent candidates (from translation):
+
 \begin{equation}

### 2015-12-10 paper/paper.tex

--- 2015-12-10 paper/paper.tex
@@
 
-We maximize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm (Byrd et al., 1995).
+We optimize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm.
 
@@
 large is 160k
-full is 878k
+full is 878k TODO
 

### 2015-12-10 paper/paper.tex
fin chains

--- 2015-12-10 paper/paper.tex
@@
 \begin{enumerate}[label=(\arabic*)]
-    \item undergo no change (\emph{bike} $\rightarrow$ \emph{bikes})
-    \item repeat a character (\emph{star} $\rightarrow$ \emph{starring})
-    \item delete a character (\emph{believe} $\rightarrow$ \emph{believing})
-    \item modify a character (\emph{parry} $\rightarrow$ \emph{parried}).
+    \item Undergo no change (\emph{bike} $\rightarrow$ \emph{bikes})
+    \item Repeat a character (\emph{star} $\rightarrow$ \emph{starring})
+    \item Delete a character (\emph{believe} $\rightarrow$ \emph{believing})
+    \item Modify a character (\emph{parry} $\rightarrow$ \emph{parried})
 \end{enumerate}
@@
 Based on the correct segmentation of a word, we construct a probable chain.
+In addition to the split points of the segmentation, we have access to tags for each segment which give the original form of the segment.
+This information is used to construct the type (\emph{Repeat}, \emph{Modify}, etc) of the transformation.
+
 Assuming no infixes, each parent must be formed by removing segments at either the beginning or the end of the word.
@@
 
-Given a segmentation, we have two candidates for the segmentation of the parent word:
+Given a segmentation for a word which is not a base word, we have two candidates for the segmentation of the parent word:
 \begin{enumerate}[label=(\arabic*)]
-    \item
+    \item The parent is the child without a prefix.
+    \item The parent is the child without a suffix.
 \end{enumerate}
 
+We use a heuristic to decide between them.
+Our heuristic captures properties like the frequency of the parent word in the wordlist, the cosine similarity between the parent and child, and the length of the parent relative to the child.
+We select the parent candidate with the higher score.
+
+By recursing on the parent, we can generate more word-parent pairs that we want our model to generate.
+
 \subsection{Semi-supervised Model}
-In addition to maximizing the log-likelihood of the known wordlist,
-we can also use explicit feedback regarding the correct segmentation of known words.
-From these known segmentations, we can use these segmentations to reconstruct
-the morphological chains that would lead to the correct segmentation of the word.
+Using the word-parent pairs generated from the pre-segmented list, 
+in addition to maximizing the log-likelihood of the known wordlist,
+we can also maximize the log-likelihood of the correct segmentation of the training data.
 
-In addition to the neighborhood log-likelihood that we maximize,
-we also want to maximize the log-likelihood of the correct segmentation of the training data.
 If a given word can be expressed as a morphological chain, the likelihood of its correct segmentation
 is equal to the product of the likelihoods of each word-to-parent transition within the morphological chain.
-
 Therefore, the likelihood of the segmentation of the training data is the product of the likelihoods
@@
 
-We want to maximize the likelihood of the document $D$, where $w$ is a word in $D$ and $c$ is
-its segmentation:
+Let $MC(w)$ be the set of word-parent pairs constructed from the morphological chains of the pre-segmented list.
+We want to maximize the likelihood of the document $D$ in the following:
 
@@
 \end{equation}
-where $MC(w)$ represents the parents and transitions in the morphological chain of $w$ going up to the root,
-constructed using its segmentation.
+where $w$ is a word in $D$ and $c$ is its segmentation.
 

### 2015-12-10 paper/paper.tex
Paper stuff

--- 2015-12-10 paper/paper.tex
@@
 
-We use the method of Contrastive Estimation \citep{Smith05} and substitute the space of all strings $\Sigma^*$ with neighbors of each word, $N(w)$.
+We use the method of Contrastive Estimation (Smith and Eisner, 2005) and substitute the space of all strings $\Sigma^*$ with neighbors of each word, $N(w)$.
 Toward this end, we transpose pairs of consecutive letters of $w$ near both ends of the word.
@@
 
-We optimize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm.
+We maximize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm (Byrd et al., 1995).
 
@@
 
-Additionally, we extend our feature set with features regarding the translation link between
-the word and potential parent, to include information such as whether or not each candidate was
+Additionally, we extend our feature set to include whether or not each candidate was
 generated by a Turkish translation (a strong indicator for parent-ness) and how confident
 the Turkish model was in selecting a parent.
-
-Therefore, our model becomes:
-\begin{equation}
-    \begin{split}
-        P(w) &= \sum_{z \in C'(w)} P(w,z) \\
-             &= \frac{\sum_{z \in C'(w)} e^{\theta \cdot \phi(w, z, T(w), T(z))}}{\sum_{w' \in N(w)} \sum_{z \in C'(w')} e^{\theta \cdot \phi(w', z', T(w'), T(z'))}}
-    \end{split}
-\end{equation}
-where $C'(w)$ is the potential parents of $w$ updated with potential irregular translations,
-and $T(w)$ is the set of Turkish translations of $w$.
 
@@
 large is 160k
-full is 878k TODO
+full is 878k
 
@@
                      & \citeauthor{Narasimhan15} & 0.807 & 0.722 & 0.762 \\ 
-                     & Unsupervised (Baseline) & -  & - & - \\ 
-            English & Semi-supervised & - & - & - \\ 
-                    & Two language unsupervised & - & - & - \\ 
-                    & Two language semi-supervised & - & - & - \\ \hline
+                     & Unsupervised (Baseline) & 0.664  & 0.649 & 0.656 \\ 
+            English & Semi-supervised ($\lambda=15$) & 0.892 & 0.762 & 0.822 \\ 
+                    & Two language unsupervised & 0.432 & 0.652 & 0.520 \\ 
+                    & Two language semi-supervised & 0.449 & 0.759 & 0.564 \\ \hline
                     & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 \\ 
-                    & Unsupervised (Baseline) & - & - & - \\ 
-            Turkish & Semi-supervised & - & - & -  \\ 
-                    & Two language unsupervised  & - & - & -  \\ 
-                    & Two language semi-supervised  & - & - & -  \\ \hline
+                    & Unsupervised (Baseline) & 0.675 & 0.371 & 0.479 \\ 
+            Turkish & Semi-supervised ($\lambda=50$) & 0.763 & 0.642 & 0.697 \\ 
         \end{tabular}
@@
             \hline
-                    & - & - & - & - & -\\ 
+                    & müzisyenlerinden & müzisyen/ler/i/nden & - & - & -\\ 
             Turkish & - & - & - & - & -\\ 
@@
 
-We found that adding the additional Turkish candidates and features regarding translation
-did not help accuracy of training very much; this is likely due to the incompleteness of the translation
-dictionary only being able to provide potential translations for very few words, and due to the ambiguity
-of translation, it's very hard to tell whether or not something would make a good parent of a word after
-translating the word into Turkish, taking the parent in Turkish, and translating it back.
+We found that adding the additional Turkish candidates and features were actually detrimental to the 
+unsupervised case
+
 \bibliography{paper}

### 2015-12-10 paper/paper.tex
paper stuff

--- 2015-12-10 paper/paper.tex
@@
     we investigate the effects of various types of extensions to this algorithm.
-    We began by adding certain features to the model and heuristics for prefix/suffix selection.
-    In addition, we extend the unsupervised model with the log-likelihood of known word segmentations,
+    We extend the unsupervised model with the log-likelihood of known word segmentations,
     producing a semi-supervised model.
     Finally, we also attempt to use Turkish morphological parents and an English-Turkish translation
-    dictionary to help detect irregular English word segmentations and transformations.
+    dictionary to help detect and disambiguate English word segmentations and transformations.
 \end{abstract}
@@
 \section{Introduction}
-The division of a word into morphemes--basic units of meaning--- encodes information about
+The division of a word into morphemes--basic units of meaning---encodes information about
 something
 When performing morphological analysis, there are a few different cases to consider: words that share similar
-meaning but appear in diffrent syntactic contexts (``German" ---``Germanic"), new words generated from
+meaning but appear in diffrent syntactic contexts (``German"---``Germanic"), new words generated from
 the combination of two separate words (``bookcase"|``book"---``case"), words with dissimilar meaning but
@@
 \begin{enumerate}[label=(\arabic*)]
-    \item Undergo no change (\emph{bike} $\rightarrow$ \emph{bikes})
-    \item Repeat a character (\emph{star} $\rightarrow$ \emph{starring})
-    \item Delete a character (\emph{believe} $\rightarrow$ \emph{believing})
-    \item Modify a character (\emph{parry} $\rightarrow$ \emph{parried})
+    \item undergo no change (\emph{bike} $\rightarrow$ \emph{bikes})
+    \item repeat a character (\emph{star} $\rightarrow$ \emph{starring})
+    \item delete a character (\emph{believe} $\rightarrow$ \emph{believing})
+    \item modify a character (\emph{parry} $\rightarrow$ \emph{parried}).
 \end{enumerate}
@@
 Based on the correct segmentation of a word, we construct a probable chain.
-In addition to the split points of the segmentation, we have access to tags for each segment which give the original form of the segment.
-This information is used to construct the type (\emph{Repeat}, \emph{Modify}, etc) of the transformation.
-
 Assuming no infixes, each parent must be formed by removing segments at either the beginning or the end of the word.
@@
 
-Given a segmentation for a word which is not a base word, we have two candidates for the segmentation of the parent word:
+Given a segmentation, we have two candidates for the segmentation of the parent word:
 \begin{enumerate}[label=(\arabic*)]
-    \item The parent is the child without a prefix.
-    \item The parent is the child without a suffix.
+    \item
 \end{enumerate}
 
-We use a heuristic to decide between them.
-Our heuristic captures properties like the frequency of the parent word in the wordlist, the cosine similarity between the parent and child, and the length of the parent relative to the child.
-We select the parent candidate with the higher score.
+\subsection{Semi-supervised Model}
+In addition to maximizing the log-likelihood of the known wordlist,
+we can also use explicit feedback regarding the correct segmentation of known words.
+From these known segmentations, we can use these segmentations to reconstruct
+the morphological chains that would lead to the correct segmentation of the word.
 
-By recursing on the parent, we can generate more word-parent pairs that we want our model to generate.
-
-\subsection{Semi-supervised Model}
-Using the word-parent pairs generated from the pre-segmented list, 
-in addition to maximizing the log-likelihood of the known wordlist,
-we can also maximize the log-likelihood of the correct segmentation of the training data.
-
+In addition to the neighborhood log-likelihood that we maximize,
+we also want to maximize the log-likelihood of the correct segmentation of the training data.
 If a given word can be expressed as a morphological chain, the likelihood of its correct segmentation
 is equal to the product of the likelihoods of each word-to-parent transition within the morphological chain.
+
 Therefore, the likelihood of the segmentation of the training data is the product of the likelihoods
@@
 
-Let $MC(w)$ be the set of word-parent pairs constructed from the morphological chains of the pre-segmented list.
-We want to maximize the likelihood of the document $D$ in the following:
+We want to maximize the likelihood of the document $D$, where $w$ is a word in $D$ and $c$ is
+its segmentation:
 
@@
 \end{equation}
-where $w$ is a word in $D$ and $c$ is its segmentation.
+where $MC(w)$ represents the parents and transitions in the morphological chain of $w$ going up to the root,
+constructed using its segmentation.
 
 We can simplify this product by letting $TR(D)$ be all pairs of words and parents implied in each of
-the morphological chains of all words in $D$ ---
-then, the log-likelihood of our training data can be expressed as
+the morphological chains of all words in $D$---then, the log-likelihood of our training data can be expressed as
 
@@
 from a morphological parser developed at Boğaziçi University. The Turkish wordlist
-was obtained from a significantly smaller corpus --- at 1 million words, compared
+was obtained from a significantly smaller corpus---at 1 million words, compared
 to the English corpus of 18 million.
-
-We tested on 4 different sizes of wordvectors:
-of wordvector 
-small is 2.5k
-med is 
-large is 160k
-full is 878k
 
@@
         \end{tabular}
-        \caption{Accuracies of Various Models}
+        \caption{Accuracies of Various Models. For our model, we used the top 48000 English words in the wordlist,
+        and the MorphoChallenge 2010 Train/Development sets for training and testing, respectively.}
     \end{center}
@@
             \hline
-                    & - & - & - & - & -\\ 
-            Turkish & - & - & - & - & -\\ 
-                    & - & - & - & - & -\\ 
-                    & - & - & - & - & -\\ \hline
-                    & - & - & - & -  & -\\ 
-                    & - & - & - & -  & -\\ 
-    English (dual-language) & - & - & - & -  & -\\ 
-                    & - & - & - & -  & -\\ 
-                    & - & - & - & -  & -\\ \hline
+                    & kompozisyonudur & kompozisyon/u/dur & aklInIz & aklI/nIz & akl/In/Iz\\ 
+            Turkish & ayet & ayet & genCliGinizi & genC/liGi/ni/zi & genC/liG/iniz/i\\ 
+                    & rUyalarInIz & rUya/lar/InIz & irdeleyin & irdeley/in & irdele/yin\\ 
+                    & borular & boru/lar & indirgenemez & indirgenemez & indirge/n/emez\\ \hline
                 \end{tabular}
@@
 performance, from an accuracy of 65.5\% to 81.7\% with English, and from 47.9\% to 70.3\% with Turkish.
+This is a substantial but expected increase; observing the unsupervised model reveals that it performs
+many single-letter
 
-We found that adding the additional Turkish candidates and features were actually detrimental to the 
-unsupervised case
-
+We found that adding the additional Turkish candidates and features regarding translation
+did not help accuracy of training very much; this is likely due to the incompleteness of the translation
+dictionary only being able to provide potential translations for very few words, and due to the ambiguity
+of translation, it's very hard to tell whether or not something would make a good parent of a word after
+translating the word into Turkish, taking the parent in Turkish, and translating it back.
 \bibliography{paper}

### 2015-12-10 paper/paper.tex
stuff

@@
                     & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 \\ 
-                    & Unsupervised (Baseline) & 0.675 & 0.371 & 0.479 \\ 
-            Turkish & Semi-supervised ($\lambda=50$) & 0.763 & 0.642 & 0.697 \\ 
+            Turkish & Unsupervised (Baseline) & 0.675 & 0.371 & 0.479 \\ 
+                    & Semi-supervised ($\lambda=50$) & 0.763 & 0.642 & 0.697 \\ \hline
         \end{tabular}
@@
 We found that utilizing information from the gold segmentations, we were able to substantially increase our
-performance, from an accuracy of 65.5\% to 81.7\% with English, and from 47.9\% to 70.3\% with Turkish.
+performance, from an accuracy of 65.5\% to 81.7\% with English, and from 47.9\% to 69.7\% with Turkish.
 This is a substantial but expected increase; observing the unsupervised model reveals that it performs
-many single-letter
+many segmentations with nonexistent affixes, which are rectified by providing input about known affixes.
 

### 2015-12-10 paper/paper.tex
plots

@@
 \usepackage{url,hyperref}
+\usepackage{tikz,pgfplots}
+
+\newcommand\engsmallw{2514}
+\newcommand\engmedw{48146}
+\newcommand\englargew{126537}
+\newcommand\engfullw{878036}
+\newcommand\tursmallw{1592}
+\newcommand\turmedw{14909}
+\newcommand\turlargew{89023}
+\newcommand\turfullw{617298}
+
+\newcommand\engsmallv{10000}
+\newcommand\engmedv{100000}
+\newcommand\englargev{500000}
+\newcommand\engfullv{1477761}
+\newcommand\tursmallv{10000}
+\newcommand\turmedv{100000}
+\newcommand\turlargev{300000}
+\newcommand\turfullv{593836}
 
@@
 \section{Results}
+\begin{figure}
+    \begin{tikzpicture}
+        \begin{axis}[
+                height=0.8\linewidth,
+                width=\linewidth,
+                xmin=1000, xmax=300000,
+                ymin=0, ymax=1,
+                xmode=log,
+                xlabel=Training Vocab Size,
+                ylabel=F-1 Score,
+            ]
+            \addplot coordinates {
+                (\engsmallw, 0.556)
+                (\engmedw, 0.656)
+                (\englargew, 0.383)
+            };
+            \addlegendentry{English}
+            \addplot coordinates {
+                (\tursmallw, 0.124)
+                (\turmedw, 0.479)
+                (\turlargew, 0.520)
+            };
+            \addlegendentry{Turkish}
+        \end{axis}
+    \end{tikzpicture}
+    \label{fig:unsupervised}
+    \caption{Performance of unsupervised model after frequency thresholding training data}
+\end{figure}
+
+\begin{figure}
+    \begin{tikzpicture}
+        \begin{axis}[
+                height=0.8\linewidth,
+                width=\linewidth,
+                xmin=0, xmax=20,
+                ymin=0, ymax=1,
+                xlabel=$\alpha$,
+                ylabel=F-1 Score,
+                legend pos=south east,
+            ]
+            \addplot coordinates {
+                (0, 0.656)
+                (0.1, 0.683)
+                (1, 0.752)
+                (3, 0.803)
+                (5, 0.815)
+                (8, 0.824)
+                (10, 0.823)
+                (12, 0.821)
+                (15, 0.822)
+                (20, 0.817)
+            };
+            \addlegendentry{English}
+            \addplot coordinates {
+                (0, 0.479)
+                (0.1, 0.580)
+                (1, 0.663)
+                (3, 0.668)
+                (5, 0.670)
+                (8, 0.680)
+                (10, 0.680)
+                (20, 0.691)
+                (50, 0.697)
+            };
+            \addlegendentry{Turkish}
+        \end{axis}
+    \end{tikzpicture}
+    \label{fig:supervised}
+    \caption{Performance of semi-supervised model based on $\alpha$.}
+\end{figure}
+
 We found that utilizing information from the gold segmentations, we were able to substantially increase our

### 2015-12-10 paper/paper.tex
papery stuff

@@
 \section{Introduction}
-The division of a word into morphemes--basic units of meaning---encodes information about
-something
+The division of words into morphemes--basic units of meaning, such as root words,
+affixes, et cetera---encodes information about how words are constructed and derive their meaning,
+as well as how languages evolve.
+
 When performing morphological analysis, there are a few different cases to consider: words that share similar

### 2015-12-10 paper/paper.tex
more paper stuff

@@
 as well as gold standard segmentation training and development sets of approximately 
-1000 randomly selected words. Gold standard segmentation is based on the CELEX database.
+1000 randomly selected words. The MorphoChallenge 2010 gold standard segmentation is based on the CELEX database.
 For Turkish, the word list contains approximately 670,000 words, with gold standards
@@
 was obtained from a significantly smaller corpus---at 1 million words, compared
-to the English corpus of 18 million.
+to the English corpus of 18 million. For a word-word translation dictionary,
+we used a English-Turkish dictionary
+(retrieved from \url{http://www.fen.bilkent.edu.tr/~aykutlu/sozluk.txt}).
 
-\subsection{Evaluation Model}
+\subsection{Evaluation}
+Performance was measured by looking at the segmentation points within each word
+(i.e. the points within the word where it is split into separate morphemes)
+and evaluating the F1 measure, i.e. the harmonic mean of precision and recall;
+precision is the fraction of the segmentation points that are present
+in the gold segmentation; recall is the fraction of the gold segmentation points
+that are correctly identified by our estimator.
 

### 2015-12-10 paper/paper.tex
cite

@@
 % TODO wordvector citation?
-It has been shown that words can be mapped to word vectors in a reasonably-sized dimensional space such that semantically similar words align in similar directions.
-Thus we can compute a \emph{cosine similarity} as a distance metric between words which captures semantic similarity.
+It has been shown that words can be mapped to word vectors in a reasonably-sized dimensional space such that semantically similar words align in similar directions \citep{Mikolov13}, thus we can compute a \emph{cosine similarity} as a distance metric between words which captures semantic similarity.
 Given vector representations of two words, the cosine similarity is measured as the ratio of the dot product between the word vectors and the product of the norms of the word vectors.
 
-In our work, we generate 200-dimensional vectors for words by using Word2Vec, which looks at word alignment  and co-occurrence patterns across many documents.
+In our work, we generate 200-dimensional vectors for words using Word2Vec, which looks at word alignment and co-occurrence patterns across many documents.
 In generating our vectors, we use text from Wikipedia,
-to generate the English word vectors, along with the BOUN corpus of news articles to generate Turkish word vectors.
+to generate the English word vectors, along with the BOUN Web Corpus (consisting of news articles) to generate Turkish word vectors.
 

### 2015-12-10 paper/paper.tex
comment on alpha effect

@@
 
+It should be noted that as $\alpha$ increases, the likelihood of the labeled training data matters more and more,
+and the score is positively influenced up to a point---as \alpha increases further, the labeled data
+dominates the loss function, and the information from the unstructured model ceases to be used,
+causing a slow decline in the score.
+
 We found that adding the additional Turkish candidates and features regarding translation

### 2015-12-10 paper/paper.tex
oops

--- 2015-12-10 paper/paper.tex
@@
 \documentclass[11pt,twocolumn]{article} 
+\usepackage[utf8]{inputenc}
 \usepackage{simpleConference}
@@
 
-\author{Calvin Huang (calvinh@mit.edu)\\
-    Brian Shimanuki (bshimanuki@mit.edu)\\
-    Charlotte Chen (czchen@mit.edu)
-\\
+\author{Calvin Huang (calvinh@mit.edu), 6.867\\
+    Brian Shimanuki (bshimanuki@mit.edu), 6.867\\
+    Charlotte Chen (czchen@mit.edu), 6.806\\
 }
@@
 
+It should be noted that as $\alpha$ increases, the likelihood of the labeled training data matters more and more,
+and the score is positively influenced up to a point---as $\alpha$ increases further, the labeled data
+dominates the loss function, and the information from the unstructured model ceases to be used,
+causing a slow decline in the score.
+
 We found that adding the additional Turkish candidates and features regarding translation
 did not help accuracy of training very much; this is likely due to the incompleteness of the translation
-dictionary only being able to provide potential translations for very few words, and due to the ambiguity
+dictionary only being able to provide potential translations for very few words. Furthermore, due to the ambiguity
 of translation, it's very hard to tell whether or not something would make a good parent of a word after
-translating the word into Turkish, taking the parent in Turkish, and translating it back.
+translating the word into Turkish, taking the parent in Turkish, and translating it back. The additional candidates,
+for the most part, reduced accuracy by increasing the number of incorrect parent choices,
+as opposed to their intended result of providing parents that our candidate generation algorithms missed.
 \bibliography{paper}

### 2015-12-10 paper/paper.tex
stuff

--- 2015-12-10 paper/paper.tex
@@
 \section{Introduction}
-The division of words into morphemes--basic units of meaning, such as root words,
-affixes, et cetera---encodes information about how words are constructed and derive their meaning,
-as well as how languages evolve.
+The division of a word into morphemes--smallest units with meaning---encodes information about both the structure
+of the word as well as its semantic meaning. Most work in the field looks at either orthographic information or
+semantic information. \citet{Narasimhan15} integrates these approaches by representing a word
+with a series of derivations from the base form of the word, thereby capturing both types of information in each derivation.
+At each step, the similarity across the derivation can be analyzed using both character similarity measures and semantic similarity measures.
 
 When performing morphological analysis, there are a few different cases to consider: words that share similar
-meaning but appear in diffrent syntactic contexts (``German"---``Germanic"), new words generated from
+meaning but appear in different syntactic contexts (``German"---``Germanic"), new words generated from
 the combination of two separate words (``bookcase"|``book"---``case"), words with dissimilar meaning but
-derived from one another ``insuffcient"---``sufficient"). In the first and last case, words are often related
-by the addition of preffixes, suffixes, and other affixes. This allows us to form morphological chains - which
+derived from one another ``insufficient"---``sufficient"). In the first and last case, words are often related
+by the addition of prefixes, suffixes, and other affixes. This allows us to form morphological chains - which
 link a base word to its derivative through a series of affixations.
-We wish to
 
@@
 \section{Unsupervised Model}
-Our model builds on work done by \citet{Narasimhan15} in constructing \emph{morphological chains} to model the morphological segmentation of words in a language in an unsupervised manner.
+Our model builds on work by \citeauthor{Narasimhan15} in constructing \emph{morphological chains} to model the morphological segmentation of words in a language in an unsupervised manner.
 The idea behind morphological chains is that complex words are constructed by attaching morphemes simpler words.
@@
 % TODO wordvector citation?
-It has been shown that words can be mapped to word vectors in a reasonably-sized dimensional space such that semantically similar words align in similar directions \citep{Mikolov13}, thus we can compute a \emph{cosine similarity} as a distance metric between words which captures semantic similarity.
+It has been shown that words can be mapped to word vectors in a reasonably-sized dimensional space such that semantically similar words align in similar directions.
+Thus we can compute a \emph{cosine similarity} as a distance metric between words which captures semantic similarity.
 Given vector representations of two words, the cosine similarity is measured as the ratio of the dot product between the word vectors and the product of the norms of the word vectors.
 
-In our work, we generate 200-dimensional vectors for words using Word2Vec, which looks at word alignment and co-occurrence patterns across many documents.
+In our work, we generate 200-dimensional vectors for words by using Word2Vec, which looks at word alignment  and co-occurrence patterns across many documents.
 In generating our vectors, we use text from Wikipedia,
-to generate the English word vectors, along with the BOUN Web Corpus (consisting of news articles) to generate Turkish word vectors.
+to generate the English word vectors, along with the BOUN corpus of news articles to generate Turkish word vectors.
 
@@
 We approach this by trying to maximize the likelihood of observing the words in $D$ from the space of all constructible strings from the alphabet, $\Sigma^*$.
-We use the EM-algorithm to iteratively maximize the likelihood over $\theta$ as given by
+We maximize the likelihood over $\theta$ as given by
 
@@
 
-Our EM-algorithm will then operate on the log-likelihood, which takes the following form:
+Our algorithm operates on the log-likelihood, which takes the following form:
 
@@
 \subsection{Prediction}
-The EM-algorithm yields an optimal $\theta^*$ of weights of the features generated from $D$.
+The algorithm yields an optimal $\theta^*$ of weights of the features generated from $D$.
 Then given a test word $w$, we can predict the probability of a given parent candidate $z$ by computing $P(z|w)=e^{\theta^*\cdot\phi(w,z)}$ as in \autoref{equ:prob_par}.
@@
 as well as gold standard segmentation training and development sets of approximately 
-1000 randomly selected words. The MorphoChallenge 2010 gold standard segmentation is based on the CELEX database.
+1000 randomly selected words. Gold standard segmentation is based on the CELEX database.
 For Turkish, the word list contains approximately 670,000 words, with gold standards
@@
 was obtained from a significantly smaller corpus---at 1 million words, compared
-to the English corpus of 18 million. For a word-word translation dictionary,
-we used a English-Turkish dictionary
-(retrieved from \url{http://www.fen.bilkent.edu.tr/~aykutlu/sozluk.txt}).
+to the English corpus of 18 million.
 
-\subsection{Evaluation}
-Performance was measured by looking at the segmentation points within each word
-(i.e. the points within the word where it is split into separate morphemes)
-and evaluating the F1 measure, i.e. the harmonic mean of precision and recall;
-precision is the fraction of the segmentation points that are present
-in the gold segmentation; recall is the fraction of the gold segmentation points
-that are correctly identified by our estimator.
+The Turkish language contains characters like \c{c} and \u{g} which are not in the English alphabet.
+In computing with them and in our results below, we have preprocessed the Turkish texts to replace characters not found in English with a capitalized letter.
+(We convert all words to lowercase beforehand, so there are no ambiguous symbols.)
+
+\subsection{Evaluation Model}
 
@@
     \label{fig:supervised}
-    \caption{Performance of semi-supervised model based on $\alpha$.}
+    \caption{Performance of the semi-supervised model based on $\alpha$.}
 \end{figure}

### 2015-12-10 paper/paper.tex

--- 2015-12-10 paper/paper.tex
@@
 \usepackage{simpleConference}
+\usepackage[margin=1.125in]{geometry}
 \usepackage{natbib}
@@
 When performing morphological analysis, there are a few different cases to consider: words that share similar
-meaning but appear in diffrent syntactic contexts (``German"---``Germanic"), new words generated from
+meaning but appear in different syntactic contexts (``German"---``Germanic"), new words generated from
 the combination of two separate words (``bookcase"|``book"---``case"), words with dissimilar meaning but
-derived from one another ``insuffcient"---``sufficient"). In the first and last case, words are often related
-by the addition of preffixes, suffixes, and other affixes. This allows us to form morphological chains - which
+derived from one another ``insufficient"---``sufficient"). In the first and last case, words are often related
+by the addition of prefixes, suffixes, and other affixes. This allows us to form morphological chains - which
 link a base word to its derivative through a series of affixations.
-We wish to
+
+Most work in the field looks at either orthographic information or
+semantic information. \citet{Narasimhan15} integrates these approaches by representing a word
+with a series of derivations from the base form of the word, thereby capturing both types of information in each derivation.
+At each step, the similarity across the derivation can be analyzed using both character similarity measures and semantic similarity measures.
 
@@
 was obtained from a significantly smaller corpus---at 1 million words, compared
-to the English corpus of 18 million. For a word-word translation dictionary,
-we used a English-Turkish dictionary
+to the English corpus of 18 million.
+For a word-word translation dictionary, we used a English-Turkish dictionary
 (retrieved from \url{http://www.fen.bilkent.edu.tr/~aykutlu/sozluk.txt}).
+
+The Turkish language contains characters like \c{c} and \u{g} which are not in the English alphabet.
+In computing with them and in our results below, we have preprocessed the Turkish texts to replace characters not found in English with a capitalized letter.
+(We convert all words to lowercase beforehand, so there are no ambiguous symbols.)
 

### 2015-12-10 paper/paper.tex
sutf

@@
 
+To build the two-language model, we first train the Turkish model with semi-supervised data.
+Using this model and an English-Turkish/Turkish-English dictionary, we derive an extended
+set of possible English parent candidates for English words, and additional features. 
+These are then used in conjunction with training the English model on the unlabeled word lists,
+and also with the labeled training data.
+
 \section{Experimental Validation}
@@
                      & Unsupervised (Baseline) & 0.664  & 0.649 & 0.656 \\ 
-            English & Semi-supervised ($\lambda=15$) & 0.892 & 0.762 & 0.822 \\ 
+            English & Semi-supervised ($\alpha=15$) & 0.892 & 0.762 & 0.822 \\ 
                     & Two language unsupervised & 0.432 & 0.652 & 0.520 \\ 
@@
             Turkish & Unsupervised (Baseline) & 0.675 & 0.371 & 0.479 \\ 
-                    & Semi-supervised ($\lambda=50$) & 0.763 & 0.642 & 0.697 \\ \hline
+                    & Semi-supervised ($\alpha=50$) & 0.763 & 0.642 & 0.697 \\ \hline
         \end{tabular}

### 2015-12-10 paper/paper.tex
stuff

@@
 
+We try using differently sized training sets by varying frequency thresholds for the words.
+As training sets often contain some amount of error and noise (the English wordlist
+contains an html tag as a word), the frequency threshhold is a good approximation
+of error; words with low frequency are may be mispelling or words from other languages.
+We find that for datasets ranging from 2500 words to 48,000 words give similar quality results,
+with a sharp drop off in scores afterwards.
+
 We found that adding the additional Turkish candidates and features regarding translation

### 2015-12-10 pysrc/fileio.py
OOPS

@@ -20,7 +20,8 @@ def read_wordcounts(filename, dump = False):
     with open(filename) as f:
         for line in f:
             count, word = line.split()
-            d[word] = int(count)
+            if len(word) > 3:
+                d[word] = int(count)
     if dump:
         pickle.dump(d, open("data/wordcounts.p", "wb"))
     return d

### 2015-12-10 pysrc/morphochain.py

@@ -263,6 +263,8 @@ class MorphoChain(object):
         return parentsScores.most_common(k)
 
     def genSeg(self, word):
+        if len(word) < 3:
+            return word
         SEG_SEP = "/"
         if "-" in word:
             segmentations = []

### 2015-12-10 pysrc/fileio.py
affix vector

@@ -26,6 +26,21 @@ def read_wordcounts(filename, dump = False):
     return d
 
 
+def read_somewords(filename, dump = False):
+    """
+    Returns a dict of words and their corresponding counts.
+    """
+    d = Counter()
+    with open(filename) as f:
+        for line in f:
+            word = line.split()
+            word = word[0]
+            d[word] = 200
+    if dump:
+        pickle.dump(d, open("data/somewords.p", "wb"))
+    return d
+
+
 # Reads from morpho challege
 def readCorpus(filename, dump=None):
     f = open(filename, 'r')
@@ -74,6 +89,8 @@ def read_dictionary(filename):
                 tur_to_eng[tur].append(eng)
     return eng_to_tur, tur_to_eng
 
+
+#read_somewords('data/somewords.txt', True)
 
 if __name__=="__main__":
     for lang in ('eng', 'tur'):

### 2015-12-10 pysrc/morphochain.py

@@ -98,7 +98,7 @@ class MorphoChain(object):
             if affix in self.prefixNeighbours:
                 for n, score in self.prefixNeighbours[affix][:ParentType.TOPNEIGHBOURS]:
                     if parent + n in self.vocab:
-                        d['neighbours_COR_S'] = affix
+                        d['neighbours_COR_P'] = affix
         else:  # some sort of suffix
             if z.transformtype == ParentType.SUFFIX:
                 affix = w[lenparent:]

### 2015-12-10 pysrc/morphochain.py
wao

@@ -97,7 +97,7 @@ class MorphoChain(object):
                     # d['diffpre_' + prefix] = cos_sim
             if affix in self.prefixNeighbours:
                 for n, score in self.prefixNeighbours[affix][:ParentType.TOPNEIGHBOURS]:
-                    if parent + n in self.vocab:
+                    if n + parent in self.vocab:
                         d['neighbours_COR_S'] = affix
         else:  # some sort of suffix
             if z.transformtype == ParentType.SUFFIX:
@@ -123,7 +123,7 @@ class MorphoChain(object):
             # # affix correlation TODO check for each case
             if affix in self.suffixNeighbours:
                 for n, score in self.suffixNeighbours[affix][:ParentType.TOPNEIGHBOURS]:
-                    if w[:lenparent - len(affix)] + n in self.vocab:
+                    if w[:-len(affix)] + n in self.vocab:
                         d['neighbours_COR_S'] = affix
          # parent is not in word list
         if z.parentword not in self.vocab:

### 2015-12-15 paper/paper.tex
butterfly souffle

@@
 We approach this by trying to maximize the likelihood of observing the words in $D$ from the space of all constructible strings from the alphabet, $\Sigma^*$.
-We maximize the likelihood over $\theta$ as given by
+We maximize the log-likelihood over $\theta$ as given by
 
@@
     \begin{split}
-        L(\theta; D) &= \prod_{w^*\in D} P(w^*) \\
-        &= \prod_{w^*\in D} \sum_{z\in C(w^*)} P(w^*, z) \\
-        &= \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in\Sigma^*}\sum_{z\in C(w) e^{\theta\cdot\phi(w,z)}}} \\
+        L(\theta; D) &= \log \prod_{w^*\in D} P(w^*) \\
+        &= \log \prod_{w^*\in D} \sum_{z\in C(w^*)} P(w^*, z) \\
+        &= \log \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in\Sigma^*}\sum_{z\in C(w) e^{\theta\cdot\phi(w,z)}}} \\
     \end{split}
@@
 
-With this substitution, we can formulate the contrastive likelihood as
+With this substitution, we can formulate the contrastive log-likelihood as
 
 \begin{equation}
-    L_C(\theta; D) = \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}}
+    L_C(\theta; D) = \log \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}}
 \end{equation}
@@
     \begin{split}
-        L_C(\theta; D) &= \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}} \\
+        L_C(\theta; D) &= \log \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}} \\
         &\qquad - \lambda \|\theta\|^2
@@
 
-Our algorithm operates on the log-likelihood, which takes the following form:
+Simplifying, this becomes:
 
@@
     \begin{split}
-        L_L(\theta; D) &= \sum_{w^*\in D} \Bigg(\log \sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)} \Bigg.\\
+        L_C(\theta; D) &= \sum_{w^*\in D} \Bigg(\log \sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)} \Bigg.\\
         &\qquad\Bigg. -\log \sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)} \Bigg) \\
@@
     \begin{split}
-        &\frac{\partial L_L(\theta; D)}{\partial \theta_j} \\
+        &\frac{\partial L_C(\theta; D)}{\partial \theta_j} \\
         &= \sum_{w^*\in D} \Bigg( \frac{\sum_{z\in C(w^*)} \phi_j(w^*,z)\cdot e^{\theta\cdot\phi(w^*,z)}} {\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} \Bigg. \\
@@
 
-We optimize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm.
+We optimize $L_C(\theta; D)$ with its gradient using the LBFGS-B algorithm.
 
@@
                      & \citeauthor{Narasimhan15} & 0.807 & 0.722 & 0.762 \\ 
-                     & Unsupervised (Baseline) & 0.664  & 0.649 & 0.656 \\ 
-            English & Semi-supervised ($\alpha=15$) & 0.892 & 0.762 & 0.822 \\ 
-                    & Two language unsupervised & 0.432 & 0.652 & 0.520 \\ 
-                    & Two language semi-supervised & 0.449 & 0.759 & 0.564 \\ \hline
+                     & Unsupervised (Baseline) & 0.744  & 0.756 & 0.750 \\ 
+            English & Semi-supervised ($\alpha=15$) & 0.892 & 0.777 & 0.831 \\ 
+                    & Two language unsupervised & 0.502 & 0.682 & 0.578 \\ 
+                    & Two language semi-supervised & 0.519 & 0.789 & 0.623 \\ \hline
                     & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 \\ 
-            Turkish & Unsupervised (Baseline) & 0.675 & 0.371 & 0.479 \\ 
+            Turkish & Unsupervised (Baseline) & 0.627 & 0.478 & 0.543 \\ 
                     & Semi-supervised ($\alpha=50$) & 0.763 & 0.642 & 0.697 \\ \hline
@@
             \addplot coordinates {
-                (\engsmallw, 0.556)
-                (\engmedw, 0.656)
-                (\englargew, 0.383)
+                (\engsmallw, 0.626)
+                (\engmedw, 0.746)
+                (\englargew, 0.603)
             };
@@
                 (\tursmallw, 0.124)
-                (\turmedw, 0.479)
-                (\turlargew, 0.520)
+                (\turmedw, 0.529)
+                (\turlargew, 0.570)
             };
@@
             \addplot coordinates {
-                (0, 0.656)
-                (0.1, 0.683)
-                (1, 0.752)
+                (0, 0.716)
+                (0.1, 0.753)
+                (1, 0.782)
                 (3, 0.803)
@@
             \addplot coordinates {
-                (0, 0.479)
-                (0.1, 0.580)
+                (0, 0.549)
+                (0.1, 0.620)
                 (1, 0.663)
@@
 We found that utilizing information from the gold segmentations, we were able to substantially increase our
-performance, from an accuracy of 65.5\% to 81.7\% with English, and from 47.9\% to 69.7\% with Turkish.
+performance, from an accuracy of 75\% to 81.7\% with English, and from 54.3\% to 69.7\% with Turkish.
 This is a substantial but expected increase; observing the unsupervised model reveals that it performs

### 2015-12-15 pysrc/morphochain.py

@@ -14,6 +14,8 @@ from pysrc import accuracy
 ParentTransformation = namedtuple('ParentTransformation',
                                   ['parentword', 'transformtype', 'olangconfidence'])
 
+TOPNEIGHBOURS = 1
+
 
 class ParentType():
     STOP = 'STOP'
@@ -23,7 +25,6 @@ class ParentType():
     DELETE = 'DELETE'
     REPEAT = 'REPEAT'
     OLANG = 'OLANG'
-    TOPNEIGHBOURS = 7
 
 
 class MorphoChain(object):
@@ -96,7 +97,7 @@ class MorphoChain(object):
                     # # FIX: this is currently storing only the last prefix into d['diff']
                     # d['diffpre_' + prefix] = cos_sim
             if affix in self.prefixNeighbours:
-                for n, score in self.prefixNeighbours[affix][:ParentType.TOPNEIGHBOURS]:
+                for n, score in self.prefixNeighbours[affix][:TOPNEIGHBOURS]:
                     if n + parent in self.vocab:
                         d['neighbours_COR_P'] = affix
         else:  # some sort of suffix
@@ -122,7 +123,7 @@ class MorphoChain(object):
                     # d['diffsuff_' + suffix ] = cos_sim
             # # affix correlation TODO check for each case
             if affix in self.suffixNeighbours:
-                for n, score in self.suffixNeighbours[affix][:ParentType.TOPNEIGHBOURS]:
+                for n, score in self.suffixNeighbours[affix][:TOPNEIGHBOURS]:
                     if w[:-len(affix)] + n in self.vocab:
                         d['neighbours_COR_S'] = affix
          # parent is not in word list
@@ -300,7 +301,12 @@ class MorphoChain(object):
         elif candidate[1] == ParentType.PREFIX:
             return word[:-p_len] + SEG_SEP + self.genSeg(parent)
         else:
-            raise Exception((word, parent))
+            lencommonprefix = 0
+            for a, b in zip(word, parent):
+                if a != b:
+                    break
+                lencommonprefix += 1
+            return word[lencommonprefix:] + self.genSeg(parent)[:lencommonprefix]
 
     def similarity(self, w1, w2):
         if w1 not in self.wordvectors or w2 not in self.wordvectors:

### 2015-12-15 pysrc/morphochain.py
tofu

@@ -97,10 +97,14 @@ class MorphoChain(object):
                     # # FIX: this is currently storing only the last prefix into d['diff']
                     # d['diffpre_' + prefix] = cos_sim
             if affix in self.prefixNeighbours:
+<<<<<<< Updated upstream
                 for n, score in self.prefixNeighbours[affix][:TOPNEIGHBOURS]:
                     if n + parent in self.vocab:
                         d['neighbours_COR_P'] = affix
                 for n, score in self.prefixNeighbours[affix][:TOPNEIGHBOURS]:
+=======
+                for n, score in self.prefixNeighbours[affix][:ParentType.TOPNEIGHBOURS]:
+>>>>>>> Stashed changes
                     if n + parent in self.vocab:
                         d['neighbours_COR_P'] = affix
         else:  # some sort of suffix
